### 2023-10-27 paper/supplementary.qmd
Update paper and supplementary information

---
title: Supporting information
bibliography: bibliography.bib
format:
  html:
    embed-resources: true
    include-in-header: mathjax.html
  pdf: default
  docx: default
---

# Maud's kinetic model

## Parameters

@tbl-params shows all of Maud's unknown parameters along with their dimensions

Note that Maud’s metabolic model includes some quantities that are not treated
as parameters in its statistical model, including temperatures, compartment
volumes and the formation energy of water. Maud treats these quantities as
if they were known precisely: they can be configured by the user or default
values can be used. Although in practice there can be considerable uncertainty
regarding these quantities, we chose to disregard this uncertainty in the
interest of simplicity.

| Parameter | Modelled quantity | Dimensions | 
| --------- | ----------------- | ---------- |
| $\Delta_fG$| Formation energy | metabolites |
| $k_M$ | Michaelis Menten constants | Substrates of all enzyme/reactions and products of reversible enzyme/reactions |
| $k_I$ | Inhibition constants | Inhibiting metabolite/compartments of enzyme/reactions exhibiting competitive inhibition |
| $k_{cat}$ | Rate constants | Enzyme/reactions |
| $L_0$ | Transfer constants | Allosteric interactions |
| $e_T$ | T dissociation constants | Modifying metabolites of allosteric inhibitions |
| $e_R$ | R dissociation constants | Modifying metabolites of allosteric activations |
| $k_{cat\, pme}$ | Rate constants of phosphorylation modifying enzymes | Phosphorylation modifying enzymes |
| $v_{drain}$ | Drain fluxes | Drains, experiments |
| $Enzyme$ | Enzyme concentrations | Enzymes, experiments |
| $C_{unbalanced}$ | Unbalanced metabolite/compartment concentrations | Unbalanced metabolite/compartments, experiments |
| $C_{pme}$ | Phosphorylation modifying enzyme concentrations | Phosphorylation modifying enzymes, experiments|
| $\psi$ | Membrane potentials | Experiments |

: Parameters of Maud's statistical model {#tbl-params}

Solving the steady state problem for a given set of parameters in an experiment
yields a vector $C_{balanced}$ of balanced metabolite concentrations. These
are combined with the balanced metabolite concentrations $C_{unbalanced}$ to
produce a vector $C_{mic}$ with a concentration for each metabolite/compartment
combination.

$\Delta_fG$ parameters can optionally be fixed; this can be useful for
computational purposes, as for example to avoid estimating the formation energy
of a metabolite about which there is no available information due to it only
participating in irreversible reactions.

## Rate equations

As discussed in the main text, Maud’s kinetic model decomposes into factors
contributing to the flux in a metabolic network in an experiment as shown in
equation \eqref{eq-decomposition}. For succinctness, and since Maud's model
assumes that there are no interactions between experiments, we omit any notation
referring to experiments below. We also omit any reference to the network's
drain reactions: these are modelled as being exactly determined by the values of
the parameter vector $v_{drain}$.

\begin{equation}
F(C;\theta) = Enzyme\cdot k_{cat}\cdot Reversibility \cdot Saturation \cdot Allostery \label{eq-decomposition}
\end{equation}

The term $Enzyme$ in equation \eqref{eq-decomposition} is a vector of
non-negative real numbers representing the concentration of the enzyme
catalysing each reaction.

The term $k_{cat}$ in equation \eqref{eq-decomposition} is a vector of
non-negative real numbers representing the  amount of flux carried per unit of
saturated enzyme.

The term $Reversibility$ in equation \eqref{eq-decomposition} is a vector of
real numbers capturing the impact of thermodynamic effects on the reaction's
flux, as shown in equation \eqref{eq-reversibility}.

\begin{align} 
  Reversibility &= 1 - \exp(\frac{\Delta_{r}G + RT \cdot S^T \ln(C_{mic})}{RT}) \label{eq-reversibility} \\ 
  \Delta_{r}G &= S^{T}\Delta_{f}G + n F \psi \nonumber 
\end{align}

The terms in \eqref{eq-reversibility} have the following meanings:

- $T$ is the temperature in Kelvin (a number), 
- $R$ is the gas constant (a number), 
- $\Delta_rG$ is a vector representing the Gibbs free energy change of each reaction in standard conditions, 
- $\Delta_fG$ is a vector representing the standard condition Gibbs free energy change of each metabolite's formation reaction, or in other words each metabolite's 'formation energy'. 
- $n$ is a vector representing the number of charges transported by each reaction. 
- $F$ is the Faraday constant (a number) 
- $\psi$ is a vector representing each reaction's membrane potential (these numbers only matter for reactions that transport non-zero charge)

Note that, for reactions with zero transported charge, the thermodynamic effect
on each reaction is derived from metabolite formation energies. This formulation
is helpful because, provided that all reactions' rates are calculated from the
same formation energies, they are guaranteed to be thermodynamically consistent.

The term $n$ accounts for both the charge and the directionality. For instance,
a reaction that exports 2 protons to the extracellular space in the forward
direction would have -2 charge. If a negatively charged molecule like acetate is
exported in the forward direction, $n$ would be 1.

Note that this way of modelling the effect of transported charge does not take
into account that the concentration gradient used by the transport is that of
the dissociated molecules. Thus, this expression is only correct for ions whose
concentration can be expressed in the model only in the charged form; e.g.,
protons, $K^+$, $Na^+$, $Cl^-$, etc.

The term $Saturation$ in equation \eqref{eq-decomposition} is a vector of
non-negative real numbers representing, for each reaction, the fraction of
enzyme that is saturated, i.e. bound to one of the reaction's substrates.
To describe saturation we use equation \eqref{eq-saturation}, which is taken
from @liebermeister_modular_2010. Additionally, this term captures competitive
inhibition: as competitive inhibitor concentration increases, the saturation
denominator increases, effectively decreasing the saturation of the substrate on
the total enzyme pool. Conversely, as the substrate concentration increases this
term approaches 1.

\begin{align} 
Saturation_r &= a \cdot \text{free enzyme ratio}\label{eq-saturation} \\ 
a &= \prod_{\text{s substrate}}\frac{C_{mic}^s}{k_{M}^{rs}} \nonumber \\ 
\text{free enzyme ratio} &= \begin{cases} 
  \prod_{\text{s sustrate}} (1 + \frac{C_{mic}^s}{k_{M}^{rs}})^{S_sr} 
  + \sum_{\text{c inhibitor}}\frac{C_{mic}^c}{k_I^{rc}} & r\text{ irreversible} \\ 
  -1 
  + \prod_{\text{s sustrate}} (1 + \frac{C_{mic}^s}{k_{M}^{rs}})^{S_sr} 
     + \sum_{\text{c inhibitor}}\frac{C_{mic}^c}{k_I^{rc}} 
     + \prod_{\text{p product}} (1 + \frac{C_{mic}^p}{k_{M}^{rp}})^{S_pr}  & r\text{ reversible} 
    \end{cases} \nonumber
\end{align}

The term $Allostery$ in equation \eqref{eq-decomposition} is a vector of
non-negative numbers describing the effect of allosteric regulation on each
reaction. Allosteric regulation happens when binding to a certain molecule
changes an enzyme's shape in a way that changes its catalytic behaviour. We
use equation \eqref{eq-allostery} to describe this phenomenon, following the
generalised MWC approach described in @monod_nature_1965, @changeux_2013,
@popova_generalization_1975 and @popova_description_1979.

\begin{align} 
  Allostery_r &= \frac{1}{1 + L_0^r \cdot (\text{free enzyme ratio}_r \cdot \frac{Qtense}{Qrelaxed})^{subunits}}\label{eq-allostery} \\ 
       Qtense &= 1 + \sum_{\text{i inhibitor}} \frac{C_{mic}^i}{e_T^{ri}}\nonumber \\ 
     Qrelaxed &= 1 + \sum_{\text{a activator}} \frac{C_{mic}^a}{e_R^{ra}}\nonumber
\end{align}

The parameter $L_0$ in equation \eqref{eq-decomposition} is called the transfer
constant, and the parameter vectors $e_T$ and $e_R$ are called tense and relaxed
dissociation constants respectively.

Finally, the term $Phosphorylation$ in equation \eqref{eq-decomposition}
captures the important effect whereby enzyme activity is altered due to a
coupled process of phosphorylation and dephosphorylation. This description
achieves a similar behaviour to the MWC formalism for describing allosteric
regulation, but using the rates of phosphorylation and dephosphorylation rather
than concentrations of metabolites.

\begin{align}
Phosphorylation_r &= (\frac{\alpha}{\alpha + \beta})^{subunits} \label{eq-phosphorylation}\\
\alpha &= \sum_{\text{p phosphoylator}} k_{cat\,pme}^{p} \cdot C_{pme}^p \nonumber \\
\beta &= \sum_{\text{d dephosphoylator}} k_{cat\,pme}^{d} \cdot C_{pme}^d \nonumber 
\end{align}

# Methionine case study {#sec-methionine-case-study}

## Dataset generation

## Prior distributions compared with true parameter values

@tbl-case-study-params shows the prior distributions we used for independent
parameters.

| parameter name                      | 1% prior quantile | 99% prior quantile | true value   | prior Z-score of true value |
| ----------------------------------- | ----------------- | ------------------ | ------------ | --------------------------- |
| $e_{𝑅}^{𝐶𝐵𝑆1,𝑎𝑚𝑒𝑡𝑐}$              | 3.430e-06      | 0.002480           | 9.3e-05 | 0.004                       |
| $e_{𝑅}^{𝐺𝑁𝑀𝑇1,𝑎𝑚𝑒𝑡𝑐}$            | 3.000e-05      | 0.002000           | 2.000e-05 | -2.787                     |
| $e_{𝑅}^{𝑀𝐴𝑇3,𝑎𝑚𝑒𝑡𝑐}$              | 1.000e-04      | 0.001000           | 3.170e-04 | 0.003                       |
| $e_{𝑅}^{𝑀𝐴𝑇3,𝑚𝑒𝑡−𝐿𝑐}$             | 4.500e-04      | 0.000800           | 6.000e-04 | 0.000                       |
| $e_{𝑅}^{𝑀𝑇𝐻𝐹𝑅1,𝑎ℎ𝑐𝑦𝑠𝑐}$         | 1.120e-07      | 0.000081           | 2.000e-06 | -0.101                     |
| $e_{𝑇}^{𝐺𝑁𝑀𝑇1,𝑚𝑙𝑡ℎ𝑓𝑐}$           | 1.120e-05      | 0.008050           | 2.290e-04 | -0.136                     |
| $e_{𝑇}^{𝑀𝑇𝐻𝐹𝑅1,𝑎𝑚𝑒𝑡𝑐}$          | 1.120e-07      | 0.000081           | 1.500e-05 | 0.549306                    |
| $k_{𝑐𝑎𝑡}^{𝐴𝐻𝐶1}$                     | 1.200e+02      | 400.000000         | 2.340e+02 | 0.179861                    |
| $k_{𝑐𝑎𝑡}^{𝐵𝐻𝑀𝑇1}$                   | 6.000e+00      | 35.000000          | 1.380e+01 | -0.135                     |
| $k_{𝑐𝑎𝑡}^{𝐶𝐵𝑆1}$                     | 1.000e+01      | 188.000000         | 7.020e+00 | -2.887                     |
| $k_{𝑐𝑎𝑡}^{𝐺𝑁𝑀𝑇1}$                   | 7.000e-01      | 60.000000          | 1.050e+01 | 0.352083                    |
| $k_{𝑐𝑎𝑡}^{𝑀𝐴𝑇1}$                     | 8.200e-02      | 59.100000          | 7.900e+00 | 0.44375                     |
| $k_{𝑐𝑎𝑡}^{𝑀𝐴𝑇3}$             | 5.890e-01      | 424.000000         | 1.990e+01 | 0.080556                    |
| $k_{𝑐𝑎𝑡}^{𝑀𝐸𝑇𝐻−𝐺𝑒𝑛}$ | 4.840e-01      | 349.000000         | 1.160e+00 | -1.209                     |
| $k_{𝑐𝑎𝑡}^{𝑀𝑆1kcatMS1}$                | 1.000e+00      | 3.300000           | 1.770e+00 | -0.091                     |
| $k_{𝑐𝑎𝑡}^{𝑀𝑇𝐻𝐹𝑅1}$       | 1.300e+00      | 4.200000           | 3.170e+00 | 0.183333                    |
| $k_{𝑐𝑎𝑡}^{𝑃𝑅𝑂𝑇1}$                   | 1.590e-01      | 0.222000           | 2.650e-01 | 0.41875                     |
| $k_{𝐼}^{𝐺𝑁𝑀𝑇1,𝑎ℎ𝑐𝑦𝑠𝑐}$           | 2.000e-06      | 0.001400           | 5.300e-05 | 0.010                       |
| $k_{𝐼}^{𝑀𝐴𝑇1,𝑎𝑚𝑒𝑡𝑐}$              | 3.000e-04      | 0.000400           | 3.470e-04 | 0.014                       |
| $k_{𝐼}^{𝑀𝐸𝑇𝐻−𝐺𝑒𝑛,𝑎ℎ𝑐𝑦𝑠𝑐}$     | 1.000e-06      | 0.000030           | 6.000e-06 | 0.021                       |
| $k_{𝑀}^{𝐴𝐻𝐶1,𝑎ℎ𝑐𝑦𝑠𝑐}$             | 5.220e-05      | 0.037600           | 2.320e-05 | -2.050                     |
| $k_{𝑀}^{𝐴𝐻𝐶1,𝑎𝑑𝑛𝑐}$                | 1.670e-07      | 0.000120           | 5.660e-06 | 0.081944                    |
| $k_{𝑀}^{𝐴𝐻𝐶1,ℎ𝑐𝑦𝑠−𝐿𝑐}$            | 1.580e-07      | 0.000114           | 1.060e-05 | 0.318056                    |
| $k_{𝑀}^{𝐵𝐻𝑀𝑇1,ℎ𝑐𝑦𝑠−𝐿𝑐}$          | 1.200e-05      | 0.000032           | 1.980e-05 | 0.049                       |
| $k_{𝑀}^{𝐵𝐻𝑀𝑇1,𝑔𝑙𝑦𝑏𝑐}$            | 4.720e-05      | 0.034000           | 8.460e-03 | 0.659028                    |
| $k_{𝑀}^{𝐶𝐵𝑆1,ℎ𝑐𝑦𝑠−𝐿𝑐}$            | 1.000e-06      | 0.000025           | 4.240e-05 | 3.090                       |
| $k_{𝑀}^{𝐶𝐵𝑆1,𝑠𝑒𝑟−𝐿𝑐}$             | 2.000e-06      | 0.000004           | 2.830e-06 | 0.004                       |
| $k_{𝑀}^{𝐺𝑁𝑀𝑇1,𝑎𝑚𝑒𝑡𝑐}$            | 1.300e-05      | 0.009400           | 5.200e-04 | 0.1375                      |
| $k_{𝑀}^{𝐺𝑁𝑀𝑇1,𝑎ℎ𝑐𝑦𝑠𝑐}$           | 4.100e-07      | 0.000295           | 1.100e-05 | 0.000                       |
| $k_{𝑀}^{𝐺𝑁𝑀𝑇1,𝑔𝑙𝑦𝑐}$              | 5.480e-05      | 0.039500           | 2.540e-03 | 0.189583                    |
| $k_{𝑀}^{𝐺𝑁𝑀𝑇1,𝑠𝑎𝑟𝑐𝑠𝑐}$          | 3.730e-09      | 0.000003           | 1.000e-07 | 0.000                       |
| $k_{𝑀}^{𝑀𝐴𝑇1,𝑚𝑒𝑡−𝐿𝑐}$             | 1.400e-05      | 0.000720           | 1.070e-04 | 0.074                       |
| $k_{𝑀}^{𝑀𝐴𝑇1,𝑎𝑡𝑝𝑐}$                | 5.270e-05      | 0.038000           | 2.030e-03 | 0.125694                    |
| $k_{𝑀}^{𝑀𝐴𝑇3,𝑚𝑒𝑡−𝐿𝑐}$             | 4.470e-05      | 0.032200           | 1.130e-03 | -0.029                     |
| $k_{𝑀}^{𝑀𝐴𝑇3,𝑎𝑡𝑝𝑐}$                | 5.270e-05      | 0.038000           | 2.370e-03 | 0.179167                    |
| $k_{𝑀}^{𝑀𝐸𝑇𝐻−𝐺𝑒𝑛,𝑎𝑚𝑒𝑡𝑐}$      | 7.000e-06      | 0.000013           | 9.370e-06 | -0.135                     |
| $k_{𝑀}^{𝑀𝑆1,5𝑚𝑡ℎ𝑓𝑐}$                | 3.320e-06      | 0.002390           | 6.940e-05 | -0.124                     |
| $k_{𝑀}^{𝑀𝑆1,ℎ𝑐𝑦𝑠−𝐿𝑐}$              | 1.000e-06      | 0.000003           | 1.710e-06 | -0.054                     |
| $k_{𝑀}^{𝑀𝑇𝐻𝐹𝑅1,𝑚𝑙𝑡ℎ𝑓𝑐}$         | 7.500e-05      | 0.000088           | 8.080e-05 | -0.158                     |
| $k_{𝑀}^{𝑀𝑇𝐻𝐹𝑅1,𝑛𝑎𝑑𝑝ℎ𝑐}$         | 1.600e-05      | 0.000028           | 2.090e-05 | -0.105                     |
| $k_{𝑀}^{𝑃𝑅𝑂𝑇1,𝑚𝑒𝑡−𝐿𝑐}$           | 4.500e-05      | 0.000085           | 4.390e-05 | -2.507                     |
| $𝐿_{0}^{𝐶𝐵𝑆1}$                          | 3.730e-02      | 26.800000          | 1.030e+00 | 0.017                       |
| $𝐿_{0}^{𝐺𝑁𝑀𝑇1}$                        | 3.730e-02      | 26.800000          | 1.310e+02 | 0.3875                      |
| $𝐿_{0}^{𝑀𝐴𝑇3}$                          | 3.730e-03      | 2.680000           | 1.080e-01 | 0.037                       |
| $𝐿_{0}^{𝑀𝑇𝐻𝐹𝑅1}$                      | 1.120e-01      | 80.500000          | 3.920e-01 | -1.018                     |

: Parameter specification, marginal prior distributions and true parameter values used in our case study. {#tbl-case-study-params}

$\Delta_fG$ parameters for most metabolites were fixed; those that were
modelled as unknown had a multivariate normal prior distribution derived from
eQuilibrator [@beberEQuilibratorPlatformEstimation2021].

The values for $\Delta_fG$ parameters, as well as all other model parameters,
can be found by inspecting the file `priors.toml` which is online at <https://github.com/biosustain/Methionine_model/blob/main/data/methionine/priors.toml>.

The results of all reported Maud runs can be found at <https://github.com/biosustain/Methionine_model/blob/main/results>.

# References

::: {#refs}
:::