### 2023-10-28 paper/main.qmd
Update paper and supplementary, add main .gitignore

---
title: "Bayesian regression modelling facilitates quantitative modelling of cell metabolism"
keywords: Bayesian inference, kinetic models of cell metabolism 
abstract: >
  This paper presents Maud, a command-line application that implements
  Bayesian statistical inference for kinetic models of biochemical metabolic
  reaction networks. Maud takes into account quantitative information from omics
  experiments and background knowledge, as well as structural information about
  kinetic mechanisms, regulatory interactions and enzyme knockouts. Below, we
  review the existing options in this area, explain how Maud improves on the state
  of the art, describe the intended modelling workflow and illustrate its use with
  an example application.
author: 
  - name: "Teddy Groves"
    orcid: 0000-0002-7109-3270
    email: tedgro@biosustain.dtu.dk
    affiliations: 
    - id: biosustain
      name: DTU Biosustain
  - name: "Nicholas Luke Cowie"
    affiliations: 
    - ref: biosustain
  - name: "Lars Keld Nielsen"
    affiliations: 
    - ref: biosustain
bibliography: bibliography.bib
format:
  acs-pdf:
    keep-tex: true
    journal:
      id: asbcd6  #This is an identifier for the target journal: from https://pubs.acs.org/, the identifier is the part of the URL after https://pubs.acs.org/journal/, e.g. for J. Org. Chem. it would be joceah.
      type: article  #article, note, communication, review, letter, or perspectivee
      layout: traditional #traditional or twocolumn, twocolumn currently broken for longtables
      hide-email: false # if enabled, hides email addresses (and fax/phone)
      abbreviations: IR,NMR,UV
  acs-html:
   include-in-header: mathjax.html
   embed-resources: true
  docx: default
---

# Introduction

A kinetic model of cellular metabolism aims to express what is known
about a cellular process in the form of an *in silico* representation
of the underlying network of chemical reactions. Kinetic models can be
used to improve production of target molecules, determine regulatory
networks [@christodoulou_reserve_2018] and identify potential drug targets
[@deberardinis_fundamentals_2016;@Liberti2017]. However, the use of kinetic
models in practice is hindered by their dependence on noisy and uncertain
information sources. Quantitative in vivo measurements of chemical abundances,
and in vitro measurements relating to kinetic parameters, both contain vital
information but are notoriously inaccurate (REFERENCES). Practically useful
kinetic modelling therefore requires a principled statistical approach that
encompasses multiple possible model parameterisations.

Bayesian statistical inference can combine the structural information implicit
in kinetic models with knowledge about metabolic parameters and information from
omics measurements [@saa_construction_2016; @gopalakrishnan_k-fit_2020]. However,
kinetic models pose serious computational challenges for Bayesian inference
[@gutenkunst_2007; @raue_identifiability_2010].

The scope of a kinetic model is defined by a stoichiometric matrix, $S$, in
which rows represent metabolites, columns represent reactions, and matrix
elements $s_{ij}$ represent the stoichiometric coefficient of metabolite $i$ in
reaction $j$. The change in metabolite concentrations is:

\begin{equation}\label{eq-1}
\frac{dC}{dt} = S\cdot v - \mu\cdot C 
\end{equation}

In equation \eqref{eq-1}, C represents a vector of metabolite concentrations,
$v$ is a vector of reaction rates, and $\mu$ is the growth rate. The second term
represents the dilution due to cell growth.

In a kinetic model, the rates, $v$, are expressed as a function of the enzyme
concentrations, $E$, the metabolite concentrations, $C$, and a set of parameters,
$\theta$ as shown in equation \eqref{eq-2}

\begin{equation}\label{eq-2}
v = f(C, E, \theta)
\end{equation}

The parameters must include sufficient boundary concentrations and fluxes to
solve \eqref{eq-1}.

It is common to assume pseudo-steady state for metabolites, i.e., the rate
of fluxes towards any metabolite is much greater than the rate of change in
concentration, $𝑣 \gg \frac{𝑑𝐶}{𝑑𝑡}$. Moreover, the dilution effect is assumed
minimal, $\mu\cdot C \ll \vec{v}$ (true unless the concentration is very high).
Finally, the enzyme concentration is assumed constant for the period considered
and hence part of the parameters.

Given these assumptions and a set of values for \theta, a set of steady state
metabolite concentrations and fluxes can be found by solving for $C$ the
algebraic equation:

\begin{equation}\label{eq-3}
S\cdot f(C;\theta) = 0
\end{equation}

In a fermentation context, \eqref{eq-3} captures the rapid kinetics inside
the cell, while another set of ODEs would be used to describe the external
substrate and product concentrations, which could act as boundary parameters
to \eqref{eq-3}.

In the context of kinetic modelling, Bayesian inference is appealing because
it allows uncertainty to be represented appropriately without sacrificing
mechanistic accuracy. Measurement uncertainty can naturally be represented in a
Bayesian measurement model, whereas the prior model can represent quantitative
uncertainty about kinetic parameters. Finally, kinetic rate laws can be
represented in Bayesian data generation models with arbitrarily high fidelity.
See @gelmanBayesianDataAnalysis2020a for more about Bayesian inference and
@gelmanBayesianWorkflow2020 for a discussion of practical Bayesian workflow.

Another advantage is that Bayesian inference problems are well-posed even when
not all parameters are strongly identified. Sloppy models in which measurable
quantities are sensitive to combinations of parameters but not to individual
marginal parameter values are ubiquitous in models of biological systems
[@gutenkunst_2007; @white_2016]. The parameter correlation structure represents
the set of potential models that describe the observed data. As we demonstrate
in our case study below, capturing this correlation structure is difficult
outside a Bayesian context.

Previous Bayesian kinetic models have either sacrificed mechanistic accuracy
or have attempted to fit realistic kinetic models using obsolete or unreliable
computational methods.

The most popular algorithm for fitting Bayesian statistical models is
Markov Chain Monte Carlo (MCMC). Modern MCMC algorithms allow exploration of
high- dimensional posterior distributions, have robust failure diagnostics
[@vehtariRankNormalizationFoldingLocalization2021] and can incorporate fast
numerical solvers, thereby making inference feasible for Bayesian kinetic
models. Nonetheless, the kinetic modelling literature reports an aversion
to MCMC, rooted mainly in concerns about sampling time and the presumed
difficulty of implementing the required statistical model [@raue_joining_2013;
@saa_construction_2016]. We are only aware of two recent attempts to implement
a Bayesian kinetic modelling approach using MCMC. @stapor_pesto_2018 fitted
detailed kinetic models using relatively inefficient MCMC algorithms that
do not scale well to high dimensional parameter spaces limiting the scope
of modelling. Conversely, @st.johnBayesianInferenceMetabolic2018 utilises an
efficient sampling algorithm but uses approximate kinetics, namely lin-log
kinetics @visser_dynamic_2003, limiting the scope of interpreting parameters and
inferring cellular behaviour in experimental conditions outside the reference
dataset.

There have also been efforts to implement Bayesian inference for kinetic models
without the use of MCMC. Examples of alternative inference methods include
variational inference as in @st.johnBayesianInferenceMetabolic2018, rejection
sampling and approximate Bayesian computation @saa_construction_2016 and Laplace
approximation, in which the Fisher information matrix is used to calculate a
normal approximation around the maximum a posteriori parameter configuration
[@liebermeister_model_2021; @gopalakrishnan_k-fit_2020; @stapor_pesto_2018;
@raue_joining_2013]. Non-MCMC-based Bayesian kinetic models have limited
utility because they lack reliable diagnostic tools for verifying that their
results approximate the target posterior distribution. This is a problem because
realistic kinetic models tend to induce highly correlated, non-Gaussian, joint
probability distributions [@gutenkunst_2007; @stapor_pesto_2018].

Our application Maud is the first Bayesian kinetic model to combine biologically
realistic mechanistic accuracy---including accurate rate laws, post-
translational modification and thermodynamics---with fast, robust MCMC sampling
using adaptive Hamiltonian Monte Carlo. Further, Maud is a general-purpose
application that can be used to fit a wide range of Bayesian kinetic models.

# Results and Discussion

To demonstrate our application’s capabilities, we used Maud to analyse an
artificial dataset based on the human methionine cycle. We first used Maud
to generate simulated training and validation measurements based on plausible
parameter values, then performed posterior sampling. Next, we used Maud to
predict the validation data.

To show that Maud is robust to missing measurements we compared the results
of fitting the full dataset with an intentionally incomplete dataset. To
demonstrate why a full Bayesian approach is preferable to an approach based on a
Laplace approximation of the posterior distribution, we also fit our methionine
cycle model using the latter method and compared the results with MCMC sampling.

Finally, we investigated our results to find out what our model learned about
the contributions of different regulatory factors to the flux through GNMT, an
important reaction. This analysis illustrates how Maud can be used to generate
actionable insights about metabolism without the need for further statistical
analysis.

The methionine cycle, illustrated in @fig-methionine-reactions, is a fundamental
pathway in human metabolism, whose intermediate metabolites participate in a
variety of mechanisms which must compete for the same resources. Due to this
competition, as well as the fact that all the functions occur simultaneously,
the methionine cycle is highly regulated, with 6 known allosteric effectors
(REFERENCE). This complex regulation means that quantitative modelling of the
methionine cycle requires a detailed kinetic model: this is why we chose it as a
case study for Maud.

::: {#fig-methionine-reactions layout-ncol=1}

![](./figures/methionine-reactions.png)

The methionine cycle as modelled, with the solid black lines representing the
reactions, the green lines representing allosteric interaction, and the red
lines representing allosteric inhibition. The bold fonts are the reaction names
and the regular font represents the metabolites.

:::

## Dataset and model specification

The simulated dataset and underlying kinetic model that we used for our analysis
can be found at <https://github.com/biosustain/Methionine_model/tree/main/data/methionine>
and is described in supplementary information section [REFERENCE].

We constructed a kinetic model of the methionine cycle in Maud’s format using
the description in @korendyaseva_allosteric_2008. The ordinary differential
equation system describing this model is shown in Equation \eqref{eq-meth-ode}.

\begin{align}
\frac{d[met]}{dt} &= v_{Influx} - v_{PROT} - v_{MAT} +v_{MS} + v_{BHMT} \label{eq-meth-ode} \\
\frac{d[amet]}{dt} &= v_{MAT} - v_{GNMT} - v_{METH} \nonumber \\
\frac{d[ahyc]}{dt} &= v_{GNMT} + v_{METH} - v_{AHC} \nonumber \\
\frac{d[hyc-L]}{dt} &= v_{AHC} - v_{CBS} - v_{MS} - v_{BHMT} \nonumber \\
\frac{d[5mthf]}{dt} &= v_{MTHFR} - v_{MS} \nonumber 
\end{align}

After specifying the qualitative aspects of the kinetic model, we selected
parameter values to use as ground truth by Monte Carlo sampling using a previous
model of the methionine cycle as a starting point (see @saa_general_2015 for
this model).

We used these parameters to simulate steady states in a range of plausible
experimental conditions, again using @saa_general_2015 as a starting point.
These steady states were then used to generate simulated measurements using the
measurement model described below in @sec-methionine-measurement-model.

### Measurement model {#sec-methionine-measurement-model}

Generating an artificial dataset required a specification of the true
measurement error. For enzyme and metabolite concentration measurements
we specified a standard deviation of 0.1 on natural logarithmic scale,
corresponding to approximately 10% measurement error. For each reaction
measurement a measurement standard deviation of approximately 10% of the
simulated value.

These measurement error specifications are somewhat optimistic considering the
many sources of variation and uncertainty affecting quantitative proteomics,
metabolomics and fluxomics analyses, but are a reasonable first approximation to
a realistic set of measurements.

For our main model run, we assumed that all metabolite and enzyme concentrations
were measured, and that there was a reaction measurement for each of the
network’s elementary flux modes.

### Training/validation split

The training testing split was selected to achieve a large difference between
the fluxes of the training and testing dataset. The split was determined as
we are interested in showing how our model can fit to varied conditions, and
conditions closer to the training set are likely to be predicted well without
necessarily learning the system.

### Additional dataset with missing measurements

To gain insight into our model’s robustness to missing measurements, we
also performed a model run with the same 6 experimental datasets, but with
measurements of the metabolite S-Adenosyl-L-homocysteine, or “ahcys” removed.
Since ahcys regulates three enzymes in the methionine cycle, including one
enzyme which is also thermodynamically regulated, we expected the removal of
these measurements to yield interesting results.

### Maud input specification

We constructed inputs in Maud’s format for each of the analysed datasets, based
on the scenario that the true kinetic model was known except for parameter
values, which needed to be inferred from the training data and priors. These
inputs can be found at <https://github.com/biosustain/Methionine_model/tree/main/data>.

The prior distributions and corresponding true parameter values used in our
case study are shown in supplementary materials section [REFERENCE]. The first
two columns show the 1% and 99% quantiles of each marginal prior distribution.
True parameter value are shown in column three, and the last column shows the
z-score on log scale of the true parameter value according the marginal prior
distribution. As can be seen from the table, there are 7 parameters for which
the true value is outside the 1%-99% range. This is desirable, making the case
study more realistic, because extreme deviance from the prior distribution is
likely to occur in practice due to in vivo to in vitro measurement differences.

### Computation

We conducted adaptive Hamiltonian Monte Carlo sampling for the full and missing-
-data datasets, obtaining XXX post-warmup samples after XXX of total computation
time. For the Laplace approximation comparison, we used Maud’s Laplace mode.
Full details as well as instructions for reproducing our analysis can be found
at XXX.

## Findings

### Posterior inference

Running standard diagnostic checks indicated that the samples we generated
were from the target posterior distribution. The improved $\hat{R}$ statistic
[@vehtariRankNormalizationFoldingLocalization2021] for every sampled variable
was within 1% of 1, indicating appropriate mixing within and between Markov
chains. Additionally, the number of effective samples was high, indicating
that we generated enough posterior samples to support inferences about the
bulks of the distributions of the sampled parameters. Furthermore, we observed
no post warm-up divergent transitions, indicating that the sampler was able to
transform the log-posterior distribution, avoiding any regions with excessive
curvature that might inhibit exploration via HMC.

Posterior predictive checking indicated that our model achieved a good fit
to the simulated reaction and metabolite concentration measurements, as shown
by the graphs in the top row of figure 3. Note that the fit was good for both
training and validation measurements.

Analysis of the posterior distributions for kinetic parameters indicated that
these are highly correlated. The marginal posterior distributions for most
kinetic parameters did not shrink significantly compared with the corresponding
marginal prior distributions, even though these parameters’ joint posterior
distribution contained enough information to make accurate out of sample
predictions. In some cases, there were two-dimensional correlations such
as the one shown in the bottom left of figure 3; in this case the marginal
distribution of the two parameters is roughly banana shaped. More commonly,
however, two-dimensional pair plots were insufficient to reveal the underlying
correlation structure, as seen in the bottom-right plot in figure 3. This does
not mean that the parameters were uncorrelated, but rather that the correlations
involve more than two parameters.

Overall, our results show that Maud can fit a realistic pathway-sized dataset.
This was achieved without fixing the marginal values of kinetic parameters: the
information required to make good predictions was contained in the correlation
structure of the joint posterior distribution. This finding is consistent with
previous analyses of biological systems that found they are “sloppy”, that is,
sensitive to parameter combinations rather than marginal parameter values, with
important combinations, scales and regions of sensitivity being difficult to
ascertain in advance [@gutenkunst_2007; @poirier_revising_1998].

The question naturally arises whether the crucial high-dimensional parameter
correlations are linear or non-linear. This is relevant to the question of
model performance, as linear correlations are easier to correct for. A linearly
correlated posterior space would also be easier to summarise. We address this
question below in @sec-laplace.

This case study illustrates the type of kinetic model and dataset that Maud
can fit. The model we analysed has XXX reactions, XXX state variables and XXX
parameters. Posterior sampling with adaptive Hamiltonian Monte Carlo generated
XXX ESS/hour. Generalising from this result, we conclude that it is feasible
to use this method to fit models in the same order of magnitude, but not,
for example, genome-scale kinetic models. To fit larger models, faster steady
state solving methods or alternative inference algorithms will be required.
@sec-laplace addresses whether Laplace approximation is a suitable candidate.

### Comparison with Laplace approximation {#sec-laplace}

We found that the Laplace method was not able to produce an accurate posterior
approximation for our model and dataset.

Our overall strategy was to compare samples generated using MCMC, which we
were confident could be treated as draws from the true posterior distribution,
with approximate posterior samples generated using the Laplace method. We were
unable to generate approximate posterior samples for our main methionine cycle
case study using Laplace approximation, as the algorithm could not recover from
solver failures caused by unrealistic parameter configurations. We therefore
made a comparison for a simpler model. This input can be found at <https://github.com/biosustain/Methionine_model/tree/main/data/example_ode>.

MCMC sampling for our comparison input yielded XXX samples in XXX time; Laplace
sampling yielded XXX samples in XXX time. The diagnostics indicated that our
algorithm was able to find the maximum a posteriori parameter configuration,
approximate the Hessian and use these quantities to generate approximate
posterior samples. The results can be found at <https://github.com/biosustain/Methionine_model/tree/main/results>.

@fig-laplace summarises the results of comparing the samples generated using
each method. As can be seen from the top left plot, the Laplace method does not
provide a good approximation to the true posterior distribution in this case,
as the marginal distribution of the total log probability density is clearly
different. This impression was confirmed using the Kolmogorov-Smirnov test,
which is a test to differentiate two empirical univariate distributions. The
two distributions were significantly different with a p-value indistinguishable
from zero.

The difference between the Laplace approximation output and the true posterior
distribution manifests not only in the parameter space, but also in the
measurement space. Figure @fig-laplace frame B compares the 5%-95% interval
for flux measurement log likelihoods in the true posterior with the Laplace
approximation; lower log likelihood values indicate that the modelled and
measured values are further away. The graph shows that the Laplace approximation
yielded significantly worse predictions than the true posterior, even for the
training data.

To further explore why this is the case we compared samples from the true
posterior and the Laplace approximation for the pairwise marginal distributions
of two Michaelis-Menten constants [XXX]- and [XXX]: see the bottom right cell of
figure @fig-laplace. This comparison demonstrates that the Laplace method is not
able to capture the correct relationships between parameters’ distributions.

This result shows that MCMC, while slower than Laplace approximation, is
unfortunately preferable for posterior inference in this case. We expect that
the Laplace method will produce worse approximations the more complex the target
model. Since the approximation is already unacceptable for our simple test
model, we recommend that Maud users use MCMC sampling in preference to Laplace
approximation if possible when fitting realistic Bayesian kinetic models.

Our results here also provide circumstantial evidence that the parameter
correlations in Bayesian kinetic model posteriors tend to be non-linear, as
a posterior with only linear correlations would likely be more germane to
Laplace approximation. A conclusion that we drew from this analysis was that
the results of fitting our model cannot be summarised simply, for example by
fitting a multivariate normal distribution to the posterior draws. We therefore
recommend that Maud users store the full set of MCMC draws rather than using
such an approximation. This does not preclude the possibility that there is
an alternative, more compact, way to summarise the results of Bayesian kinetic
model inference; we leave research into this topic to future work.

::: {#fig-laplace layout-ncol=1}

![](./figures/laplace.png)

CAPTION: FILL THIS IN!

:::

### Effect of missing metabolite concentration measurements

Comparing model runs with and without the ahcys measurements showed that Maud
can produce sensible results even from incomplete metabolomics data.

@fig-missing shows that, as might be expected, the model with missing
measurements did not correctly infer the missing ahcys concentrations.
Nonetheless, the remaining measured metabolites were still well predicted,
suggesting that information about the network is still preserved despite the
missing measurements. Comparison of flux measurements in both models also
indicated that removing the ahcys measurement did not result in catastrophic
model failure.

The missing measurements did affect Maud’s ability to infer parameter values
correctly. The lower left plot of @fig-missing shows that the model with full
metabolomics learned the true value for the displayed dissociation constant,
despite this value being far from the mean of the corresponding marginal prior
distribution. In contrast, the model with missing measurements stayed in the
neighbourhood of the prior.

This result is reassuring because not having access to all measurements is a
common situation in multi-omics studies. For instance, measuring all metabolites
in a pathway can be infeasible because of limitations of mass spectrometers,
availability of standards, column effects, and compartmentalisation. However,
provided that sufficient information is available from other sources, our
approach can produce sensible results from incomplete metabolomics data.

::: {#fig-missing layout-ncol=1}

![](./figures/missing.png)

Results of removing concentration measurements for the metabolite $ahcys_c$
from our case study dataset.  (**Top**) A comparison of metabolite concentration
residuals between the full measurement dataset (blue-green) and the missing-data
dataset (grey), displayed on natural logarithmic scale. The missing-data
model was unable to predict the withheld $ahcys_c$ concentrations.  (**Bottom
Left**) The marginal posterior distribution for the Michaelis constant
$K_m^{AHC1,ahcys_c}$ in each model, alongside the true parameter value used to
generate both datasets. The true value is recovered by the complete-data model
but not by the missing-data model.  (**Bottom Right**) The distribution of total
log-likelihood for out-of-sample flux measurements in both models. There is a
significant overlap between the two distributions, suggesting that removing the
$ahcys_c$ measurement did not cause catastrophic prediction failure. However,
overall the complete-data model tended to make better predictions.

:::

### Application to regulatory understanding

To demonstrate how Maud's output can be used to yield useful metabolic insights
we used the results of our case study to explain why the flux of the enzyme
$GNMT$ is higher in condition 1 than in condition 2. GNMT is an irreversible
enzyme that is homotropically activated by its substrate, competitively
inhibited by its product and heterotropically inhibited. The diverse regulation
makes it the ideal test case to elucidate regulatory changes.

@fig-decomposition shows the regulatory description of GNMT, according to the
results our main case study analysis. Each curve shows the marginal posterior
distribution of the ratio of the corresponding regulatory component in condition
1 compared with condition 2. A positive value indicates that the component
was increased in condition 1 relative to condition 2, with 0 indicating no
difference. The probability, according to our model, of the component acting in
each direction is given by the relative area under the curve on each side of the
0 point.

Our model correctly inferred that saturation and allosteric effects were the
main drivers of regulation between the two conditions in this case, with the
curves for each component aligning with the ground truth shown in red.

Importantly, this form of analysis takes into account all modelled sources of
uncertainty, including uncertainty about the true values of the flux in each
condition, and propagates this uncertainty into the final conclusion. Our result
shows that Maud could be used in this realistic case not only to provide a user
with an explanation for an observed difference in fluxes, but also a reasonable
judgement as to the explanation's robustness.

::: {#fig-decomposition layout-ncol=1}

![](./figures/decomposition.png)

Illustration of how analysing a system with Maud can yield actionable insights
about the underlying metabolic network.  (**Top Left**) A schematic of the
regulatory interactions associated with the enzyme $GNMT1$. Dashed green
lines represents allosteric activation, dashed red lines indicate allosteric
inhibition and solid red lines represent competitive inhibition. (**Bottom
Left**) Comparison of marginal posterior distributions for $GNMT$ flux in
conditions 1 and 2. (**Right**) Log-scale ratios of the regulatory elements
defined in equation \eqref{eq-decomposition}. Note that the $Reversibility$
and $k_cat$ components are excluded: this is because this reaction was modelled
as irreversible, and $k_{cat}$ was modelled as constant across conditions.
These plots identify why flux in condition 2 is higher than in condition 1: the
flux increase is due to allostery and saturation with no control from enzyme
concentration changes.

:::

# Methods

Maud is a command line application implementing Bayesian inference for
a wide range of realistic kinetic models. Maud is written in Python
[@vanrossumPythonReferenceManual2009], designed for use on Windows, macOS
and Linux, registered on the Python Package Index as `maud-metabolic-models`,
documented at <https://maud-metabolic-models.readthedocs.io> and actively
developed and maintained at <https://github.com/biosustain/Maud/>.

To use Maud, a user must first collate appropriate input information, represent
it in files with Maud’s required formats (see section (REFERENCE)below).  Maud’s
command line interface provides commands for inference, simulation and making
out-of-sample predictions. Results are stored in files, using a structured,
interoperable format.

## Input format

Maud inputs are structured directories, somewhat inspired by the PEtab
format [@SchmiesterSch2021]. A Maud input directory must contain a toml
[@preston-wernertomandgedampradyunTOMLSpecification0rc2020] file called
`config.toml` which gives the input a name, configures how Maud will be run
and tells Maud where to find the other files, allowing these to have custom
names. It must also include a file containing a kinetic model definition,
a file specifying information about parameters and a file with information
experiments. The required structure of these files is documented at <https://
maud-metabolic-models.readthedocs.io/en/latest/inputting.html>. The input is
validated against a Pydantic [@pydanticdevelopersPydantic2022] data model which can be
inspected at <https://github.com/biosustain/Maud/tree/main/maud/data_model>.

We chose to implement a custom input format despite the existence of standard
formats in similar areas, including SBML [@keatingSBMLLevelExtensible2020]
and PEtab [@SchmiesterSch2021]. This choice was partly motivated by the need
to ensure flexibility as Maud was developed, but there are also features of
SBML and PEtab that make them structurally unsuitable in this context. Our
requirements for an input format included that it be mathematics-free, so that
all mathematical details are encapsulated in source code, and that it has a
detailed, verifiable structure. These requirements made toml more attractive
than SBML: toml is easier for humans to read and edit and can straightforwardly
be validated using tools like Pydantic. Further, an SBML representation of our
desired input would not contain differential equations.  It would therefore not
be interoperable with most SBML targeting software, which typically assumes that
differential equations are available and does not know about Maud’s structure.

## Kinetic model

Maud’s kinetic model decomposes into factors as shown in equation
\eqref{eq-decomposition}.

\begin{equation}
F(C;\theta) = Enzyme\cdot k_{cat}\cdot Reversibility \cdot Saturation \cdot Allostery \label{eq-decomposition}
\end{equation}

Each of the terms on the right-hand side of \eqref{eq-decomposition} is a
function of $C$ and $\theta$. This idea is taken from @noor_note_2013. The
terms usefully gather physically meaningful and conceptually distinct factors
contributing to reaction fluxes. $Enzyme$ captures the effect of enzyme
concentration, $k_{cat}$ that of enzyme efficiency, $Reversibility$ quantifies
thermodynamic effects, $Saturation$ the effect of enzyme availability and
$Allostery$ the effect of post translational modifications.

We used the model of enzyme saturation from @liebermeister_modular_2010
and the generalised Monod-Wyman-Changeux model of Allosteric
regulation introduced in [@monod_nature_1965; @changeux_2013;
@popova_generalization_1975; @popova_description_1979] and used more recently
in @matosGRASPComputationalPlatform2022. To capture the effect on enzyme
activity of coupled phosphorylation and dephosphorylation processes we developed
a new mathematical model inspired by the generalised MWC model of allosteric
regulation. Full details of all mathematical aspects of Maud’s kinetic model can
be found in supplementary material section (REFERENCE).

## Statistical model

Maud represents information from measurements using generalised linear
regression models that probabilistically connect realised measurements with
true values of the measurable quantities. Information from other sources
is represented using a prior distribution over a set of latent parameters.
The parameters and the measurable quantities are connected by a generative
model encompassing Maud’s kinetic model as well as the steady state equation
\eqref{eq-3}. Together, the measurement model, prior model and generative model
determine a joint probability function that assigns a probability density to any
possible combination of measurements and parameters.

Below we describe the prior and measurement models, as the generative model has
already been discussed above.

### Prior model

Maud’s prior model includes unknown parameters corresponding to quantities
in the kinetic model that are assumed to be unknown, other than steady state
metabolite concentrations and fluxes, which are derived from the values of
other parameters by solving the steady state problem. See table (REFERENCE) in
this paper's supplementary materials for a description of all these parameters
and their dimensions. Note that some quantities in Maud’s kinetic model are
not treated as parameters: for example, temperatures, compartment volumes and
the formation energy of water. Maud treats these quantities as if they were
known precisely: they can be configured by the user or default values can be
used. Although in practice there can be considerable uncertainty regarding
these quantities, we chose to disregard this uncertainty in the interest of
simplicity.

Except for metabolites’ standard condition Gibbs energy changes of formation,
Maud uses independent normal prior distributions for parameters that can in
principle be both negative and positive. For parameters that are constrained to
be positive, Maud uses independent log-normal distributions. Formation energy
parameters have a multivariate normal prior distribution. Location, scale and
covariance parameters for all these prior distributions can be selected freely
by the user.

### Measurement model

Maud’s measurement model considers three types of measurement: metabolite
concentration measurements, enzyme concentration measurements and flux
measurements, represented by vectors $𝑦^{𝑐𝑜𝑛𝑐}$ $𝑦^{𝑒𝑛𝑧}$ and $𝑦^{𝑓𝑙𝑢𝑥}$
respectively.

All measurements are specific to an experimental condition; that is, a case
where the true state of the network, including knockouts, boundary conditions
and state variables as well as kinetic and thermodynamic parameters, can safely
be assumed to be the same. Maud’s statistical model allows for arbitrarily many
experimental conditions, and for any measurable quantity to be measured any
number of times in any condition.

Metabolite and enzyme measurements are intended to represent the results of
quantitative metabolomics and proteomics experiments. The likelihood functions
for such measurements are shown in equations Equation \eqref{eq-yconc} and
Equation \eqref{eq-yenz}.

\begin{align}
y_i^{conc} &\sim LN(\ln{\hat{y}_i^{conc}}, \sigma_i^{conc})\label{eq-yconc} \\
y_i^{enz} &\sim LN(\ln{\hat{y}_i^{enz}}, \sigma_i^{enz})\label{eq-yenz}
\end{align}

Both equations are log-normal generalised linear models with a standard
link function (the natural logarithm $\ln$) and known standard deviation
$\sigma_{𝑐𝑜𝑛𝑐}$. The use of this measurement model is motivated by the
consideration that concentrations are constrained to be non-negative, so the
measurement model should avoid assigning positive probability mass to negative
metabolite concentration values. In addition, we expect the precision of most
metabolomics and proteomics experiments to be roughly proportional to the value
of the true measured quantity, which supports a measurement model with constant
coefficient of variation. The measurement standard deviations $\sigma_{𝑐𝑜𝑛𝑐}$
and $\sigma_{𝑒𝑛𝑧}$ are assumed to be known exactly for simplicity; plausible
values can be elicited by considering the likely coefficient of variation of the
measuring apparatus.

Flux measurements, representing the results of quantitative fluxomics analyses,
are modelled using a likelihood function from a standard linear regression
model, as shown in equation \eqref{eq-yflux}. Flux measurements can be obtained
from the results of isotope labelling experiments using metabolic flux analysis,
for example as described in (Young 2014). When entering flux measurements,
it is important only to specify measurements for a network’s free fluxes, as
the values of some steady state fluxes in a metabolic network are constrained
by others, with the result that dependent fluxes cannot typically be measured
separately. If measurements of multiple dependent fluxes are entered,
information will inappropriately be double counted.

\begin{equation}
y_i^{flux} \sim LN(\ln{\hat{y}_i^{flux}}, \sigma_i^{flux})\label{eq-yflux}
\end{equation}

Our measurement model improves on analyses of metabolomics and proteomics
data that assume a regression model with normally distributed errors, whether
explicitly using a standard linear model or implicitly using ordinary least
squares fitting. This assumption is undesirable because it implies that the
measured quantity could in principle be negative, and assumes an additive
underlying random process, whereas multiplicative processes tend to better
describe real concentration data.

The use of independent measurement models for metabolite, enzyme and flux
measurements carries an implicit assumption that there are no systematic
correlations in the measurement errors. This choice was motivated by simplicity
- it would be better to use a model with potentially correlated measurements.
Similarly, it would be preferable to include measurement errors as model
parameters, thereby avoiding possible bias due to incorrect assessments of
measurement accuracy. However, we chose to use a simpler measurement model
to avoid the complexity and potential fitting issues that these changes would
entail.

Finally, the reader may wonder why Maud uses a linear regression model for
reaction flux measurements even though this creates the potential for erroneous
double counting and requires non-trivial upstream modelling, as intracellular
fluxes typically cannot be measured directly. Instead, fluxes are typically
inferred from isotope labelling experiments using metabolic flux analysis:
see @daiUnderstandingMetabolismFlux2017 for more about this method. Ideally
Maud’s measurement model for fluxes would extend from fluxes to the results of
potential labelling experiments, thereby removing the need for upstream analysis
and avoiding any double counting. This option has not yet been pursued, again
for the sake of simplicity.

## Implementation

Maud uses the Python library click [@clickdevelopersClickPythonComposable2022]
to implement a command line interface. The command line interface loads input
files as Python dictionaries, which are parsed using the Python library `toml`
[@pearsonTomlPythonLibrary2020] and then validated and converted into structured
`MaudInput` objects using Pydantic [@pydanticdevelopersPydantic2022]. Maud’s
statistical model is implemented in the probabilistic programming language Stan
[@carpenterStanProbabilisticProgramming2017] and accessed using the interface
cmdstanpy [@standevelopmentteamCmdStanPy2022]. For posterior sampling, Maud
uses the `MaudInput` to create a Stan input json file and obtain configuration
information for cmdstanpy, which it uses to (if necessary) generate a
model executable file and then trigger posterior sampling using adaptive
Hamiltonian Monte Carlo. When sampling is complete, Maud converts to the
output into the standard format `InferenceData` using the Python library arviz
[@kumarArviZUnifiedLibrary2019] and saves it as a json file, along with some
information for debugging.

Two details of Maud’s implementation are important to highlight: the method of
posterior sampling and how Maud solves the steady state problem \eqref{eq-3}.

### Posterior sampling

Although integrals of the joint probability model for kinetic models
are typically analytically intractable, they can be approximated
numerically using Markov Chain Monte Carlo (MCMC) and other methods.
Maud uses MCMC primarily because there exist many methods for verifying
that MCMC samples really do approximate the target probability
distribution: see @vehtariRankNormalizationFoldingLocalization2021 and
@taltsValidatingBayesianInference2018 for discussion of this point. In
addition, there are several examples of successful Bayesian kinetic modelling
projects using MCMC including @st.johnBayesianInferenceMetabolic2018 and
@xingModelingKineticsLargescale2010.

### Solving the steady state problem

In order to implement adaptive Hamiltonian Monte Carlo, Maud must
repeatedly solve the steady state problem \eqref{eq-3} and find its
gradients with respect to all parameters. To achieve this, Maud uses a
hybrid method involving two numerical solvers from the SUNDIALS suite
[@serbanCVODESSensitivityEnabledODE2005]: CVODES and IDAS. These solvers are
accessed via their interface from Stan. The hybrid method follows that proposed
by @margossianComputingSteadyStates2018, and involves numerically evolving
the ODE system for a short period of time, then using the difference between
the evolved and starting concentrations as the target for a numerical algebra
solver.

# Acknowledgements

# References

::: {#refs}
:::