### 2023-10-31 paper/cover_letter.qmd

---
title: Cover letter

---

# Manuscript title

Bayesian regression facilitates quantitative modelling of cell metabolism

# Name of the corresponding author

Teddy Groves (email: <mailto:tedgro@dtu.dk>)

# Names of the other authors

- Nicholas Cowie (email: <mailto:nicow@dtu.dk>)
- Lars Keld Nielsen (email: <mailto:lars.nielsen@uq.edu.au>)

# Why the paper is appropriate for ACS Synthetic Biology

Our paper is intended of the virtual special issue of ACS Synthetic Biology
concentrating on quantitative synthetic biology. It presents Maud, a software
application that we developed for fitting Bayesian statistical models of fluxes
through systems of metabolic reactions at steady state, enabling inference about
these systems from quantitative metabolomics, fluxomics and proteomics data.

Our software is a "quantitative technology and tool for synthetic
biology", as described in the [call for papers](https://axial.acs.org/biology-and-biological-chemistry/cfp-quantitative-synthetic-biology),
because it is intended to be used as a synthetic biology tool. For example, as
we illustrate in our paper, Maud can be used to identify drivers of regulation
through a pathway, thereby helping to identify or rule out knockout targets.

# Short lay summary

This paper presents Maud, a command-line application that implements
Bayesian statistical inference for kinetic models of biochemical metabolic
reaction networks.

Maud takes into account quantitative information from omics experiments
and background knowledge, as well as structural information about kinetic
mechanisms, regulatory interactions and enzyme knockouts.

Our paper reviews alternative options, illustrates Maud's use with two case
studies, and explains Maud's implementation and intended usage.