### 2023-11-01 paper/main.qmd
Implement Lars's paper comments

---
title: "Maud: Bayesian regression modelling facilitates quantitative modelling of cell metabolism"
keywords: Bayesian inference, kinetic models of cell metabolism, 
          multiomics integration, ordinary differential equations,
          regulatory anlaysis
abstract: >
  This paper presents Maud, a command-line application that implements Bayesian
  statistical inference for kinetic models of biochemical metabolic reaction
  networks. Maud takes into account quantitative information from omics
  experiments and background knowledge, as well as structural information about
  kinetic mechanisms, regulatory interactions and enzyme knockouts. Our paper
  reviews the existing options in this area, presents a case study illustrating
  how Maud can be used to analyse a metabolic network and explains the
  biological, statistical and computational design decisions underpinning Maud.
number-sections: true
author: 
  - name: "Teddy Groves"
    orcid: 0000-0002-7109-3270
    email: tedgro@biosustain.dtu.dk
    affiliations: 
      - id: biosustain
        name: DTU
        department: DTU Biosustain
        city: Kongens Lyngby
        country: Denmark
    attributes:
      corresponding: true
      equal-contributor: true
    note: These authors contributed equally to this research.
  - name: "Nicholas Luke Cowie"
    affiliations: 
      - id: biosustain
        name: DTU
        department: DTU Biosustain
        city: Kongens Lyngby
        country: Denmark
    attributes:
      equal-contributor: true
    note: These authors contributed equally to this research.
  - name: "Lars Keld Nielsen"
    affiliations: 
      - name: The University of Queensland
        department: Australian Institute for Bioengineering and Nanotechnology (AIBN)
        city: St Lucia
        country: Australia
      - id: biosustain
        name: DTU
        department: DTU Biosustain
        city: Kongens Lyngby
        country: Denmark
bibliography: bibliography.bib
format:
  acs-pdf:
    keep-tex: true
    journal:
      id: asbcd6  #This is an identifier for the target journal: from https://pubs.acs.org/, the identifier is the part of the URL after https://pubs.acs.org/journal/, e.g. for J. Org. Chem. it would be joceah.
      type: article  #article, note, communication, review, letter, or perspectivee
      layout: traditional #traditional or twocolumn, twocolumn currently broken for longtables
      hide-email: false # if enabled, hides email addresses (and fax/phone)
  acs-html:
   include-in-header: mathjax.html
   embed-resources: true
  docx: default
---

# Introduction

A kinetic model of cellular metabolism aims to express what is known
about a cellular process in the form of an *in silico* representation
of the underlying network of chemical reactions. Kinetic models can be
used to improve production of target molecules, determine regulatory
networks [@christodoulou_reserve_2018] and identify potential drug targets
[@deberardinis_fundamentals_2016;@Liberti2017]. However, the use of kinetic
models in practice is hindered by their dependence on noisy and uncertain
information sources. Quantitative in vivo measurements of chemical abundances,
and in vitro measurements relating to kinetic parameters, both contain vital
information but are notoriously inaccurate [@siskos2017; @tabb2010; @lu2017]. Practically useful
kinetic modelling therefore requires a principled statistical approach that
encompasses multiple possible model parameterisations.

Bayesian statistical inference can combine the structural information implicit
in kinetic models with knowledge about metabolic parameters and information from
omics measurements [@saa_construction_2016; @gopalakrishnan_k-fit_2020]. However,
kinetic models pose serious computational challenges for Bayesian inference
[@gutenkunst_2007; @raue_identifiability_2010].

The scope of a kinetic model is defined by a stoichiometric matrix, $S$, in
which rows represent metabolites, columns represent reactions, and matrix
elements $s_{ij}$ represent the stoichiometric coefficient of metabolite $i$ in
reaction $j$. The change in metabolite concentrations is:

\begin{equation}\label{eq-1}
\frac{dC}{dt} = S\cdot v - \mu\cdot C 
\end{equation}

where C represents a vector of metabolite concentrations, $v$ is a vector of
reaction rates, and $\mu$ is the growth rate. The second term represents the
dilution due to cell growth.

In a kinetic model, the rates, $v$, are expressed as a function of the enzyme
concentrations, $E$, the metabolite concentrations, $C$, and a set of parameters
$\theta$:

\begin{equation}\label{eq-2}
v = f(C, E, \theta)
\end{equation}

The parameters must include sufficient boundary concentrations and fluxes to
solve \eqref{eq-1}.

It is common to assume pseudo-steady state for metabolites, i.e., the rate
of fluxes towards any metabolite is much greater than the rate of change in
concentration, $𝑣 \gg \frac{𝑑𝐶}{𝑑𝑡}$. Moreover, the dilution effect is assumed
minimal, $\mu\cdot C \ll \vec{v}$ (true unless the concentration is very high).
Finally, the enzyme concentration is assumed constant for the period considered
and hence part of the parameters.

Given these assumptions and a set of values for \theta, a set of steady state
metabolite concentrations and fluxes can be found by solving for $C$ the
algebraic equation:

\begin{equation}\label{eq-3}
S\cdot f(C;\theta) = 0
\end{equation}

In a fermentation context, \eqref{eq-3} captures the rapid kinetics
inside the cell, while another set of ODEs would be used to describe the
external substrate and product concentrations, which could act as boundary
parameters to \eqref{eq-3}.

In the context of kinetic modelling, Bayesian inference is appealing because
it allows uncertainty to be represented appropriately without sacrificing
mechanistic accuracy. Measurement uncertainty can naturally be represented in a
Bayesian measurement model, whereas the prior model can represent quantitative
uncertainty about kinetic parameters. Finally, kinetic rate laws can be
represented in Bayesian data generation models with arbitrarily high fidelity.
See @gelmanBayesianDataAnalysis2020a for more details about Bayesian inference
and @gelmanBayesianWorkflow2020 for a discussion of practical Bayesian workflow.

Another advantage is that Bayesian inference problems are well-posed even when
not all parameters are strongly identified. Sloppy parameter models in which
measurable quantities are sensitive to combinations of parameters but not to
individual marginal parameter values are ubiquitous in models of biological
systems [@gutenkunst_2007; @white_2016]. The parameter correlation structure
represents the set of potential models that describe the observed data. As we
demonstrate later, capturing this correlation structure is difficult outside a
fully Bayesian context.

Previous Bayesian kinetic models have either sacrificed mechanistic accuracy
or have attempted to fit realistic kinetic models using obsolete or unreliable
computational methods.

The most popular algorithm for fitting Bayesian statistical models is Markov
Chain Monte Carlo (MCMC). Modern MCMC algorithms enable exploration of
high-dimensional posterior distributions, have robust failure diagnostics
[@vehtariRankNormalizationFoldingLocalization2021] and can incorporate fast
numerical solvers, thereby making inference feasible for Bayesian kinetic
models. Nonetheless, the kinetic modelling literature reports an aversion
to MCMC, rooted mainly in concerns about sampling time and the presumed
difficulty of implementing the required statistical model [@raue_joining_2013;
@saa_construction_2016]. We are only aware of two recent attempts to implement
a Bayesian kinetic modelling approach using MCMC. @stapor_pesto_2018 fitted
detailed kinetic models using relatively inefficient MCMC algorithms that
do not scale well to high dimensional parameter spaces limiting the scope
of modelling. Conversely, @st.johnBayesianInferenceMetabolic2018 utilises an
efficient sampling algorithm but uses approximate kinetics, namely lin-log
kinetics [@visser_dynamic_2003], limiting the scope of interpreting parameters and
inferring cellular behaviour in experimental conditions outside the reference
dataset.

There have also been efforts to implement Bayesian inference for kinetic models
without the use of MCMC. Examples of alternative inference methods include
variational inference [@st.johnBayesianInferenceMetabolic2018], rejection
sampling and approximate Bayesian computation [@saa_construction_2016] and
Laplace approximation, in which the Fisher information matrix is used to
calculate a normal approximation around the maximum a posteriori parameter
configuration [@liebermeister_model_2021; @gopalakrishnan_k-fit_2020;
@stapor_pesto_2018; @raue_joining_2013]. Non-MCMC-based Bayesian kinetic
models suffer from a lack of reliable diagnostic tools for verifying that their
results approximate the target posterior distribution. This is a problem because
realistic kinetic models tend to induce highly correlated, non-Gaussian, joint
probability distributions [@gutenkunst_2007; @stapor_pesto_2018].

Our application Maud provides the first Bayesian kinetic model to combine
biologically realistic mechanistic accuracy---including accurate rate laws,
post- translational modification and thermodynamics---with fast, robust
MCMC sampling using adaptive Hamiltonian Monte Carlo. Further, Maud is a
general-purpose application that can be used to fit a wide range of Bayesian
kinetic models.

# Results and Discussion

To demonstrate our application’s capabilities, we used Maud to analyse an
artificial dataset based on the human methionine cycle. We generated this
dataset using Maud, by simulating measurements for a set of training and
validation conditions based on plausible parameter values. Next, we performed
posterior inference for the training measurements and prediction for both
training and validation measurements.

We investigated Maud's sensitivity to missing measurements by comparing the
results of fitting a full dataset with an intentionally incomplete dataset. To
demonstrate why a full Bayesian approach is preferable to an approach based on a
Laplace approximation of the posterior distribution, we compared the results of
analysing a representative metabolic network using both methods.

Finally, we dug into our results to find out what our full dataset methionine
model learned about the contributions of different regulatory factors to the
flux through GNMT, an important reaction. This analysis illustrates how Maud can
be used to generate actionable insights about metabolism without the need for
further statistical analysis.

The methionine cycle, illustrated in @fig-methionine-reactions, is a fundamental
pathway in human metabolism, whose intermediate metabolites participate in a
variety of mechanisms which must compete for the same resources. Due to this
competition, as well as the fact that all the functions occur simultaneously,
the methionine cycle is highly regulated, with 6 known allosteric effectors
[@martinov2000; @nijhout2006; @korendyaseva_allosteric_2008]. This complex regulation means that quantitative modelling of the
methionine cycle requires a detailed kinetic model: this is why we chose it as a
case study for Maud.

::: {#fig-methionine-reactions layout-ncol=1}

![](./figures/methionine-reactions.png)

The methionine cycle as modelled, with the solid black lines representing the
reactions, the green lines representing allosteric interaction, and the red
lines representing allosteric inhibition. The bold fonts are the reaction names
and the regular font represents the metabolites.

:::

## Dataset and model specification

The simulated dataset and underlying kinetic model that we used for our analysis
can be found at <https://github.com/biosustain/Methionine_model/tree/main/data/methionine>
and is described in supporting information section 3.

We constructed a kinetic model of the methionine cycle in Maud’s format using
the description in @korendyaseva_allosteric_2008. The ordinary differential
equation system describing this model is

\begin{align}
\frac{d[met]}{dt} &= v_{Influx} - v_{PROT} - v_{MAT} +v_{MS} + v_{BHMT} \label{eq-meth-ode} \\
\frac{d[amet]}{dt} &= v_{MAT} - v_{GNMT} - v_{METH} \nonumber \\
\frac{d[ahyc]}{dt} &= v_{GNMT} + v_{METH} - v_{AHC} \nonumber \\
\frac{d[hyc-L]}{dt} &= v_{AHC} - v_{CBS} - v_{MS} - v_{BHMT} \nonumber \\
\frac{d[5mthf]}{dt} &= v_{MTHFR} - v_{MS} \nonumber 
\end{align}

After specifying the qualitative aspects of the kinetic model, we selected
parameter values to use as ground truth by Monte Carlo sampling using a previous
model of the methionine cycle as a starting point (see @saa_construction_2016 for
this model).

We used these parameters to simulate steady states in a range of plausible
experimental conditions, again using @saa_construction_2016 as a starting point.
These steady states were then used to generate simulated measurements using the
measurement model.

For enzyme and metabolite concentration measurements we specified a standard
deviation of 0.1 on natural logarithmic scale, corresponding to approximately
10% measurement error. For each reaction measurement a measurement standard
deviation of approximately 10% of the simulated value.

These measurement error specifications are somewhat optimistic considering the
many sources of variation and uncertainty affecting quantitative proteomics,
metabolomics and fluxomics analyses, but are a reasonable first approximation to
a realistic set of measurements.

For our main model run, we assumed that all metabolite and enzyme concentrations
were measured, and that there was a reaction measurement for each of the
network’s free fluxes.

The simulated experiments and measurements were split into a training and a
validation dataset in a way that achieved a large difference in flux between the
two categories. This was done to evaluate whether the fitted model is able to
extrapolate to conditions well outside the training dataset rather than merely
interpolating between the training data without necessarily learning the system.

We constructed inputs in Maud’s format for each of the analysed datasets, based
on the scenario that the true kinetic model was known except for parameter
values, which needed to be inferred from the training data and priors. These
inputs can be found at <https://github.com/biosustain/Methionine_model/tree/main/data>.

## Posterior inference

The prior distributions and corresponding true parameter values used in our
case study are shown in supporting information section 3.2. They were chosen to
reflect a plausible pre-experimental information state. In 7 cases, the marginal
prior distribution for a parameter disagrees with the true parameter values used
to generate the data; a similar situation is likely to occur in practice due to
*in vivo* vs *in vitro* measurement differences.

Running standard diagnostic checks indicated that the samples we generated
were from the target posterior distribution. The improved $\hat{R}$ statistic
[@vehtariRankNormalizationFoldingLocalization2021] for every variable of
interest was within 2% of 1, indicating appropriate mixing within and between
Markov chains. Additionally, the number of effective samples was high,
indicating that we generated enough posterior samples to support inferences
about the bulks of the distributions of the sampled parameters. Furthermore,
we observed no post warm-up divergent transitions, indicating that the sampler
was able to transform the log-posterior distribution, avoiding any regions with
excessive curvature that might inhibit exploration via HMC.

Posterior predictive checking indicated that our model achieved a good fit to
the simulated reaction and metabolite concentration measurements, as shown by
the graphs in the top row of @fig-posterior.

Analysis of the posterior distributions for kinetic parameters indicated
that these are highly correlated. The marginal posterior distributions
for most kinetic parameters did not shrink significantly compared to the
corresponding marginal prior distributions, even though these parameters’
joint posterior distribution contained enough information to make accurate out
of sample predictions. In some cases, there were two-dimensional correlations
such as the one shown in the bottom left of @fig-posterior; in this case the
marginal distribution of the two parameters is roughly banana shaped. More
commonly, however, two-dimensional pair plots were insufficient to reveal the
underlying correlation structure, as seen in the bottom-right plot in figure
@fig-posterior, which shows two marginally independent parameters. This kind of
marginal independence does not mean that the parameters are independent in the
full joint posterior distribution.

These results show that Maud can achieve an adequate fit to a realistic
pathway-sized dataset. This was achieved without fixing the marginal values
of kinetic parameters: the information required to make good predictions was
contained in the correlation structure of the joint posterior distribution.
This finding is consistent with previous analyses of biological systems that
found they are “sloppy”, that is, sensitive to parameter combinations rather
than marginal parameter values, with important combinations, scales and regions
of sensitivity being difficult to ascertain in advance [@gutenkunst_2007;
@poirier_revising_1998].

The question naturally arises whether the crucial high-dimensional parameter
correlations are linear or non-linear. This is relevant to the question of
model performance, as linear correlations are easier to correct for. A linearly
correlated posterior space would also be easier to summarise. We address this
question in the next section.

::: {#fig-posterior layout-ncol=1}

![](./figures/posterior.png)

Marginal posterior distributions from our main model run. (**Top left**)
Comparison of posterior predictive intervals with simulated flux measurement
values. All the flux measurements are within the predictive intervals,
indicating a good fit. (**Top right**) Comparison of posterior predictive
intervals with simulated concentration measurement values. These also show
a good fit. (**Bottom left**) Pairwise marginal posterior distribution
for two pairwise correlated parameters, namely $K_m^{AHC1,hcys-L}$ and
$K_m^{AHC1_ahcys}$. (**Bottom right**) Pairwise marginal posterior distribution
for two pairwise-uncorrelated parameters, namely $K_m^{MAT3,atp}$ and $K_{cat}^{MAT3}$.

:::

### Comparison with Laplace approximation {#sec-laplace}

This current case study illustrates the type of kinetic model and dataset that
Maud can fit. The model we analysed has 10 reactions, 5 state variables and 212
parameters. Generalising from our ability to fit this model in a reasonable time
using Maud, we expect that Maud can be used to fit realistic Bayesian models
of approximately the same size, but not, for example, genome-scale kinetic
models. To fit larger models, faster steady state solving methods or alternative
inference algorithms will be required. The Laplace approximation, in which the
Fisher information matrix is used to calculate a normal approximation around
the maximum a posteriori parameter configuration, is a popular strategy for
generating approximate posterior samples while avoiding a full MCMC approach.

Using the recently implemented Laplace approximation in Stan, we were unable to
generate approximate posterior samples for our main methionine cycle case study
using Laplace approximation, as the algorithm could not recover from solver
failures caused by unrealistic parameter configurations. This is not an inherent
issue of the Laplace transformation but highlights some of the challenges around
solving \eqref{eq-3} for real problems.

In order still to assess the potential use of the Laplace approximation, we made
a comparison for a simpler model. This input can be found at <https://github.com/biosustain/Methionine_model/tree/main/data/example_ode>.
MCMC sampling for this simpler model yielded 800 samples in 625 minutes, while
Laplace sampling yielded the same number of samples in only one minute.
The diagnostics indicated that our algorithm was able to find the maximum
a posteriori parameter configuration, approximate the Hessian and use these
quantities to generate approximate posterior samples. The results can be found
at <https://github.com/biosustain/Methionine_model/tree/main/results>.

Comparing the samples generated using each method shows that the Laplace
approximation does not provide a good approximation to the true posterior
distribution in this case (@fig-laplace). As can be seen from the top left
plot, the total log probability density is clearly different and this was
confirmed by a test of the equality of two empirical univariate distributions
(Kolmogorov-Smirnov, $p=1.7\times10^{-65}$).

The difference between the Laplace approximation output and the true posterior
distribution manifests not only in the parameter space, but also in the
measurement space for both training data (upper right) and validation data
(lower left). The lower log likelihood values indication that the modelled and
measured values are further away and that the Laplace approximation yielded
significantly worse predictions than the true posterior, even for the training
data.

To further explore why this is the case we compared samples from the true
posterior and the Laplace approximation for the pairwise marginal distributions
of two Michaelis-Menten constants $K_{m}^{A,r1}$ and $K_{cat}^{r1}$
(@fig-laplace, bottom right). This comparison demonstrates that the Laplace
method is not able to capture the correct relationships between parameters’
distributions.

This result shows that MCMC, while slower than Laplace approximation, is
required for posterior inference in this case. We expect that the Laplace method
will produce worse approximations the more complex the target model. Since the
approximation is already unacceptable for our simple test model, we recommend
that Maud users use MCMC sampling in preference to Laplace approximation if
possible when fitting realistic Bayesian kinetic models.

Our results here also provide circumstantial evidence that the parameter
correlations in Bayesian kinetic model posteriors tend to be non-linear, as
a posterior with only linear correlations would likely be more germane to
Laplace approximation. A conclusion that we drew from this analysis was that
the results of fitting our model cannot be summarised simply, for example by
fitting a multivariate normal distribution to the posterior draws. We therefore
recommend that Maud users store the full set of MCMC draws rather than using
such an approximation. This does not preclude the possibility that there is
an alternative, more compact, way to summarise the results of Bayesian kinetic
model inference; we leave research into this topic to future work.

::: {#fig-laplace layout-ncol=1}

![](./figures/laplace.png)

Graphical comparison of approximate posterior samples generated using Laplace
sampling (blue-green) with posterior samples generated using MCMC (dark
grey). (**Top left**) the two sets of samples clearly have different marginal
distributions for the overall log probability variable, indicating that the
Laplace samples do not accurately approximate the target distribution. (**Top
right**) The distribution of marginal posterior predictive log likelihood
values for training data flux measurements shows that the Laplace method
tended to yield much worse predictions compared with the true model (lower log
likelihood values indicate that the modelled and measured values are further
away). (**Bottom left**) The Laplace method also tended to produce worse flux
predictions for held-back test measurements. (**Bottom right**) The marginal
joint distribution of two parameters: $K_m^{A,rl}$ and $K_{cat}^{r1}$. The
Laplace method is not able to track the correct joint distribution for this
pair of parameters. This is unsurprising given that the target distribution
has position-dependent scales which are difficult for a linear approximation
to capture.

:::

## Effect of missing metabolite concentration measurements

To gain insight into our model’s robustness to missing measurements, we
also performed a model run with the same 6 experimental datasets, but with
measurements of the metabolite S-Adenosyl-L-homocysteine, or “ahcys” removed.
Since ahcys regulates three enzymes in the methionine cycle, including one
enzyme which is also thermodynamically regulated, we expected the removal of
these measurements to yield interesting results.

Comparing model runs with and without the ahcys measurements showed that Maud
can produce sensible results even from incomplete metabolomics data.

As might be expected, the model with missing measurements did not correctly
infer the missing ahcys concentrations (@fig-missing). Nonetheless, the
remaining measured metabolites were still well predicted, suggesting
that information about the network is still preserved despite the missing
measurements. Comparison of flux measurements in both models also indicated that
removing the ahcys measurement did not result in catastrophic model failure.

The missing measurements did affect Maud’s ability to infer parameter values
correctly (@fig-missing, lower left). The model with full metabolomics learned
the true value for the displayed dissociation constant, despite this value being
far from the mean of the corresponding marginal prior distribution. In contrast,
the model with missing measurements stayed in the neighbourhood of the prior.

This result is reassuring because not having access to all measurements is a
common situation in multi-omics studies. For instance, measuring all metabolites
in a pathway can be infeasible because of limitations of mass spectrometers,
availability of standards, column effects, and compartmentalisation. However,
provided that sufficient information is available from other sources, our
approach can produce sensible results from incomplete metabolomics data.

::: {#fig-missing layout-ncol=1}

![](./figures/missing.png)

Results of removing concentration measurements for the metabolite $ahcys_c$
from our case study dataset.  (**Top**) A comparison of metabolite concentration
residuals between the full measurement dataset (blue-green) and the missing-data
dataset (grey), displayed on natural logarithmic scale. The missing-data
model was unable to estimate the withheld $ahcys_c$ concentrations.  (**Bottom
Left**) The marginal posterior distribution for the Michaelis constant
$K_m^{AHC1,ahcys_c}$ in each model, alongside the true parameter value used to
generate both datasets. The true value is recovered by the complete-data model
but not by the missing-data model.  (**Bottom Right**) The distribution of total
log-likelihood for out-of-sample flux measurements in both models. There is a
significant overlap between the two distributions, suggesting that removing the
$ahcys_c$ measurement did not cause catastrophic prediction failure. However,
overall the complete-data model tended to make better predictions.

:::

## Application to regulatory understanding

To demonstrate how Maud's output can be used to yield useful metabolic insights
we used the results of our case study to explain why the flux of the enzyme
GNMT is higher in dataset 1 than in dataset 12 (@fig-decomposition). GNMT
is an irreversible enzyme that is homotropically activated by its substrate,
competitively inhibited by its product and heterotropically inhibited by 5,10-
methylenetetrahydrofolate (mlthf). The complex regulation makes it the ideal
test case to elucidate regulatory changes.

Regulation can be separated into enzyme abundance, allostery and saturation,
and we can plot the marginal posterior distribution of the ratio of the
corresponding regulatory component in dataset 1 compared with dataset 12
(@fig-decomposition, right panel). A positive value indicates that the component
was increased in dataset 1 relative to dataset 12, with 0 indicating no
difference. The probability, according to our model, of the component acting in
each direction is given by the relative area under the curve on each side of the
zero point.

Our model correctly inferred that saturation and allosteric effects were the
main drivers of regulation between the two datasets in this case, with the
curves for each component aligning with the ground truth shown in red.

Importantly, this form of analysis takes into account all modelled sources of
uncertainty, including uncertainty about the measured values of the flux in
each dataset. Our result shows that Maud could be used in this realistic case
not only to explain an observed difference in fluxes, but also to provide a
reasonable judgement as to the explanation's robustness.

::: {#fig-decomposition layout-ncol=1}

![](./figures/decomposition.png)

Illustration of how analysing a system with Maud can yield actionable insights
about the underlying metabolic network.  (**Top Left**) A schematic of the
regulatory interactions associated with the enzyme $GNMT1$. Dashed green
lines represents allosteric activation, dashed red lines indicate allosteric
inhibition and solid red lines represent competitive inhibition. (**Bottom
Left**) Comparison of marginal posterior distributions for $GNMT$ flux in
datasets 1 and 2. (**Right**) Log-scale ratios of the regulatory elements
defined in \eqref{eq-decomposition}. Note that the $Reversibility$
and $k_cat$ components are excluded: this is because this reaction was modelled
as irreversible, and $k_{cat}$ was modelled as constant across datasets.
These plots identify why flux in dataset 12 is higher than in dataset 1: the
flux increase is due to allostery and saturation with no control from enzyme
concentration changes.

:::

# Methods

Maud is a command line application implementing Bayesian inference
for a wide range of realistic kinetic models. Maud is written in
Python [@vanrossumPythonReferenceManual2009], designed for use
on Windows, macOS and Linux, is `pip` installable from the Python
Package Index as `maud-metabolic-models`, documented at <https://maud-metabolic-models.readthedocs.io> 
and actively developed and maintained at <https://github.com/biosustain/Maud/>.

To use Maud, a user must first collate appropriate input information, represent
it in files with Maud’s required formats (see supporting information section on
input format).  Maud’s command line interface provides commands for inference
via a range of algorithms including adaptive Hamiltonian Monte Carlo, as well as
commands for simulation and making out-of-sample predictions. Results are stored
in files, using a structured, interoperable format.

As well as parameter values, Maud also performs inference for derived quantities
including the components of the regulatory decomposition described below in
\eqref{eq-decomposition}, log likelihoods, simulated measurement values
and metabolic control analysis coefficients [@kacser_control_1973].

In the rest of this section we describe Maud's kinetic model at high level and
discuss Maud's statistical model, implementation and how it solves the crucial
steady state problem. Further details about Maud's kinetic model can be found in
the supporting information.

## Kinetic model

Maud’s kinetic model decomposes into factors:

\begin{equation}
F(C;\theta) = Enzyme\cdot k_{cat}\cdot Reversibility \cdot Saturation \cdot Allostery \label{eq-decomposition}
\end{equation}

Each of the terms on the right-hand side of \eqref{eq-decomposition} is a
function of $C$ and $\theta$. This idea is taken from @noor_note_2013. The
terms usefully gather physically meaningful and conceptually distinct factors
contributing to reaction fluxes. $Enzyme$ captures the effect of enzyme
concentration, $k_{cat}$ that of enzyme efficiency, $Reversibility$ quantifies
thermodynamic effects, $Saturation$ the effect of enzyme availability and
$Allostery$ the effect of post translational modifications.

We used the model of enzyme saturation from @liebermeister_modular_2010
and the generalised Monod-Wyman-Changeux model of Allosteric
regulation introduced in [@monod_nature_1965; @changeux_2013;
@popova_generalization_1975; @popova_description_1979] and used more recently
in @matosGRASPComputationalPlatform2022. To capture the effect on enzyme
activity of coupled phosphorylation and dephosphorylation processes we developed
a new mathematical model inspired by the generalised MWC model of allosteric
regulation. Full details of all mathematical aspects of Maud’s kinetic model can
be found in supporting information section 2.

## Statistical model

Maud represents information from measurements using generalised linear
regression models that probabilistically connect realised measurements with
true values of the measurable quantities. Information from other sources
is represented using a prior distribution over a set of latent parameters.
The parameters and the measurable quantities are connected by a generative
model encompassing Maud’s kinetic model as well as the steady state equation
\eqref{eq-3}. Together, the measurement model, prior model and generative model
determine a joint probability function that assigns a probability density to any
possible combination of measurements and parameters.

Below we describe the prior and measurement models, as the generative model has
already been discussed above.

### Prior model

Maud’s prior model includes unknown parameters corresponding to quantities
in the kinetic model that are assumed to be unknown, other than steady state
metabolite concentrations and fluxes, which are derived from the values of
other parameters by solving the steady state problem. See table 1 in
this paper's supporting information for a description of all these parameters
and their dimensions. Note that some quantities in Maud’s kinetic model are
not treated as parameters: for example, temperatures, compartment volumes and
the formation energy of water. Maud treats these quantities as if they were
known precisely: they can be configured by the user or default values can be
used. Although in practice there can be considerable uncertainty regarding
these quantities, we chose to disregard this uncertainty in the interest of
simplicity.

Except for metabolites’ standard condition Gibbs energy changes of formation,
Maud uses independent normal prior distributions for parameters that can in
principle be both negative and positive. For parameters that are constrained to
be positive, Maud uses independent log-normal distributions. Formation energy
parameters have a multivariate normal prior distribution. Location, scale and
covariance parameters for all these prior distributions can be selected freely
by the user.

### Measurement model

Maud’s measurement model considers three types of measurement: metabolite
concentration measurements, enzyme concentration measurements and flux
measurements, represented by vectors $𝑦^{𝑐𝑜𝑛𝑐}$ $𝑦^{𝑒𝑛𝑧}$ and $𝑦^{𝑓𝑙𝑢𝑥}$
respectively.

All measurements are specific to an experimental condition; that is, a case
where the true state of the network, including knockouts, boundary conditions
and state variables as well as kinetic and thermodynamic parameters, can safely
be assumed to be the same. Maud’s statistical model allows for arbitrarily many
experimental conditions, and for any measurable quantity to be measured any
number of times in any condition.

Metabolite and enzyme measurements are intended to represent the results of
quantitative metabolomics and proteomics experiments. The likelihood functions
for such measurements are:

\begin{align}
y_i^{conc} &\sim LN(\ln{\hat{y}_i^{conc}}, \sigma_i^{conc})\label{eq-yconc} \\
y_i^{enz} &\sim LN(\ln{\hat{y}_i^{enz}}, \sigma_i^{enz})\label{eq-yenz}
\end{align}

Both equations are log-normal generalised linear models with a standard
link function (the natural logarithm $\ln$) and known standard deviation
$ \sigma_{𝑐𝑜𝑛𝑐}$. The use of this measurement model is motivated by the
consideration that concentrations are constrained to be non-negative, so the
measurement model should avoid assigning positive probability mass to negative
metabolite concentration values. In addition, we expect the precision of most
metabolomics and proteomics experiments to be roughly proportional to the true
value of the measured quantity, which supports a measurement model with constant
coefficient of variation. The measurement standard deviations $\sigma_{𝑐𝑜𝑛𝑐}$
and $\sigma_{𝑒𝑛𝑧}$ are assumed to be known exactly for simplicity; plausible
values can be elicited by considering the likely coefficient of variation of the
measuring apparatus. Our measurement model improves on analyses of metabolomics
and proteomics data that assume a regression model with normally distributed
errors, whether explicitly using a standard linear model or implicitly using
ordinary least squares fitting.

Flux measurements, representing the results of quantitative fluxomics analyses,
are modelled using a likelihood function from a standard linear regression
model \eqref{eq-yflux}. Flux measurements can be obtained from isotope labelling
experiments using metabolic flux analysis, for example as described in (Young
2014). When entering flux measurements, it is important only to specify
measurements for a network’s free fluxes, as the values of some steady state
fluxes in a metabolic network are constrained by others, with the result that
dependent fluxes cannot typically be measured separately. If measurements of
multiple dependent fluxes are entered, information will inappropriately be
double counted.

\begin{equation}
y_i^{flux} \sim LN(\ln{\hat{y}_i^{flux}}, \sigma_i^{flux})\label{eq-yflux}
\end{equation}

The use of independent measurement models for metabolite, enzyme and flux
measurements carries an implicit assumption that there are no systematic
correlations in the measurement errors. This choice was motivated by simplicity
- it would be better to use a model with potentially correlated measurements.
Similarly, it would be preferable to include measurement errors as model
parameters, thereby avoiding possible bias due to incorrect assessments of
measurement accuracy. However, we chose to use a simpler measurement model
to avoid the complexity and potential fitting issues that these changes would
entail.

Finally, the reader may wonder why Maud uses a linear regression model for
reaction flux measurements even though this creates the potential for erroneous
double counting and requires non-trivial upstream modelling, as intracellular
fluxes typically cannot be measured directly. Instead, fluxes are typically
inferred from isotope labelling experiments using metabolic flux analysis:
see @daiUnderstandingMetabolismFlux2017 for more about this method. Ideally
Maud’s measurement model for fluxes would extend from fluxes to the results of
potential labelling experiments, thereby removing the need for upstream analysis
and avoiding any double counting. This option has not yet been pursued, again
for the sake of simplicity.

## Implementation

Maud uses the Python library click [@clickdevelopersClickPythonComposable2022]
to implement a command line interface. The command line interface loads
input files as Python dictionaries, which are parsed using the Python library
`toml` [@pearsonTomlPythonLibrary2020] and then validated and converted into
structured `MaudInput` objects using Pydantic [@pydanticdevelopersPydantic2022].
Maud’s statistical model is implemented in the probabilistic programming
language Stan [@carpenterStanProbabilisticProgramming2017] and accessed using
the interface cmdstanpy [@standevelopmentteamCmdStanPy2022]. For posterior
sampling, Maud uses the `MaudInput` to create an input file for Stan and
then trigger posterior sampling using adaptive Hamiltonian Monte Carlo. See
@betancourtConceptualIntroductionHamiltonian2018 for more about this algorithm.

When sampling is complete, Maud converts to the output into the standard format
`InferenceData` using the Python library arviz [@kumarArviZUnifiedLibrary2019]
and saves it as a json file, along with some information for debugging.
This format allows for easy checking of MCMC diagnostics including divergent
transitions, effective sample size and the improved $\hat{R}$ statistic proposed
in @vehtariRankNormalizationFoldingLocalization2021.

## Solving the steady state problem

Maud's parameters are connected with measurable quantities via the steady state
equation \eqref{eq-3}. Posterior sampling using adaptive Hamiltonian Monte
Carlo requires repeatedly solving \eqref{eq-3} and finding the gradients of its
solution with respect to all model parameters. This must be done numerically as
analytic solutions are not available for realistic kinetic models.

The speed at which this problem can be solved is tightly coupled with the
size and complexity of metabolic network that can practically be modelled. See
@timonenImportanceSamplingApproach2022a for more about considerations involved
in this kind of modelling.

To solve \eqref{eq-3} and find its gradients, Maud uses a hybrid
method involving two numerical solvers from the SUNDIALS suite
[@serbanCVODESSensitivityEnabledODE2005]: CVODES and IDAS via their
interface from Stan. The hybrid method follows that proposed by
@margossianComputingSteadyStates2018, and involves numerically evolving the ODE
system for a short period of time, then using the difference between the evolved
and starting concentrations as the target for a numerical algebra solver.

# Supporting information

The analyses described in this paper, as well as instructions for reproducing
our results, figures and paper, can be found at <https://biosustain/Methionine_model>.

The document `supporting_information.pdf` contains the following information:

- Description of Maud's input format.
- Description of Maud's kinetic model including rate equations as well as
parameters and their dimensions.
- Details of the procedure used to generate our case study results, including
generating the artificial case study datasets, specifying priors and carrying
out computation.
- Specification of the prior distribution used for the case studies.

# Acknowledgements

This research was funded by the Novo Nordisk Foundation (Grant numbers
NNF20CC0035580 NNF14OC0009473).

# References

::: {#refs}
:::