### 2023-11-01 paper/main.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[journal=asbcd6,manuscript=article,layout=traditional]{achemso}
\usepackage[version=3]{mhchem}
\newcommand*\mycommand[1]{\texttt{\emph{#1}}}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Maud: Bayesian regression modelling facilitates quantitative modelling of cell metabolism},
  pdfauthor={Teddy Groves; Nicholas Luke Cowie; Lars Keld Nielsen},
  pdfkeywords={Bayesian inference, kinetic models of cell metabolism,
multiomics integration, ordinary differential equations, regulatory
anlaysis},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\author{Teddy Groves}
\affiliation{DTU Biosustain, DTU, Kongens Lyngby, Denmark}

\altaffiliation{These authors contributed equally to this research.}

\email{tedgro@biosustain.dtu.dk}
\author{Nicholas Luke Cowie}
\affiliation{DTU Biosustain, DTU, Kongens Lyngby, Denmark}

\altaffiliation{These authors contributed equally to this research.}

\author{Lars Keld Nielsen}
\affiliation{Australian Institute for Bioengineering and Nanotechnology
(AIBN), The University of Queensland, St Lucia, Australia}
\affiliation{DTU Biosustain, DTU, Kongens Lyngby, Denmark}

\keywords{Bayesian inference, kinetic models of cell metabolism,
multiomics integration, ordinary differential equations, regulatory
anlaysis}

\title[]{Maud: Bayesian regression modelling facilitates quantitative
modelling of cell metabolism}
\makeatletter
\begin{document}
\maketitle
\begin{abstract}
This paper presents Maud, a command-line application that implements
Bayesian statistical inference for kinetic models of biochemical
metabolic reaction networks. Maud takes into account quantitative
information from omics experiments and background knowledge, as well as
structural information about kinetic mechanisms, regulatory interactions
and enzyme knockouts. Our paper reviews the existing options in this
area, presents a case study illustrating how Maud can be used to analyse
a metabolic network and explains the biological, statistical and
computational design decisions underpinning Maud.
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, enhanced, interior hidden, breakable, sharp corners, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt]}{\end{tcolorbox}}\fi

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

A kinetic model of cellular metabolism aims to express what is known
about a cellular process in the form of an \emph{in silico}
representation of the underlying network of chemical reactions. Kinetic
models can be used to improve production of target molecules, determine
regulatory networks \citep{christodoulou_reserve_2018} and identify
potential drug targets
\citep{deberardinis_fundamentals_2016, Liberti2017}. However, the use of
kinetic models in practice is hindered by their dependence on noisy and
uncertain information sources. Quantitative in vivo measurements of
chemical abundances, and in vitro measurements relating to kinetic
parameters, both contain vital information but are notoriously
inaccurate \citep{siskos2017, tabb2010, lu2017}. Practically useful
kinetic modelling therefore requires a principled statistical approach
that encompasses multiple possible model parameterisations.

Bayesian statistical inference can combine the structural information
implicit in kinetic models with knowledge about metabolic parameters and
information from omics measurements
\citep{saa_construction_2016, gopalakrishnan_k-fit_2020}. However,
kinetic models pose serious computational challenges for Bayesian
inference \citep{gutenkunst_2007, raue_identifiability_2010}.

The scope of a kinetic model is defined by a stoichiometric matrix,
\(S\), in which rows represent metabolites, columns represent reactions,
and matrix elements \(s_{ij}\) represent the stoichiometric coefficient
of metabolite \(i\) in reaction \(j\). The change in metabolite
concentrations is:

\begin{equation}\label{eq-1}
\frac{dC}{dt} = S\cdot v - \mu\cdot C 
\end{equation}

where C represents a vector of metabolite concentrations, \(v\) is a
vector of reaction rates, and \(\mu\) is the growth rate. The second
term represents the dilution due to cell growth.

In a kinetic model, the rates, \(v\), are expressed as a function of the
enzyme concentrations, \(E\), the metabolite concentrations, \(C\), and
a set of parameters \(\theta\):

\begin{equation}\label{eq-2}
v = f(C, E, \theta)
\end{equation}

The parameters must include sufficient boundary concentrations and
fluxes to solve \eqref{eq-1}.

It is common to assume pseudo-steady state for metabolites, i.e., the
rate of fluxes towards any metabolite is much greater than the rate of
change in concentration, \(𝑣 \gg \frac{𝑑𝐶}{𝑑𝑡}\). Moreover, the dilution
effect is assumed minimal, \(\mu\cdot C \ll \vec{v}\) (true unless the
concentration is very high). Finally, the enzyme concentration is
assumed constant for the period considered and hence part of the
parameters.

Given these assumptions and a set of values for \theta, a set of steady
state metabolite concentrations and fluxes can be found by solving for
\(C\) the algebraic equation:

\begin{equation}\label{eq-3}
S\cdot f(C;\theta) = 0
\end{equation}

In a fermentation context, \eqref{eq-3} captures the rapid kinetics
inside the cell, while another set of ODEs would be used to describe the
external substrate and product concentrations, which could act as
boundary parameters to \eqref{eq-3}.

In the context of kinetic modelling, Bayesian inference is appealing
because it allows uncertainty to be represented appropriately without
sacrificing mechanistic accuracy. Measurement uncertainty can naturally
be represented in a Bayesian measurement model, whereas the prior model
can represent quantitative uncertainty about kinetic parameters.
Finally, kinetic rate laws can be represented in Bayesian data
generation models with arbitrarily high fidelity. See
\citet{gelmanBayesianDataAnalysis2020a} for more details about Bayesian
inference and \citet{gelmanBayesianWorkflow2020} for a discussion of
practical Bayesian workflow.

Another advantage is that Bayesian inference problems are well-posed
even when not all parameters are strongly identified. Sloppy parameter
models in which measurable quantities are sensitive to combinations of
parameters but not to individual marginal parameter values are
ubiquitous in models of biological systems
\citep{gutenkunst_2007, white_2016}. The parameter correlation structure
represents the set of potential models that describe the observed data.
As we demonstrate later, capturing this correlation structure is
difficult outside a fully Bayesian context.

Previous Bayesian kinetic models have either sacrificed mechanistic
accuracy or have attempted to fit realistic kinetic models using
obsolete or unreliable computational methods.

The most popular algorithm for fitting Bayesian statistical models is
Markov Chain Monte Carlo (MCMC). Modern MCMC algorithms enable
exploration of high-dimensional posterior distributions, have robust
failure diagnostics
\citep{vehtariRankNormalizationFoldingLocalization2021} and can
incorporate fast numerical solvers, thereby making inference feasible
for Bayesian kinetic models. Nonetheless, the kinetic modelling
literature reports an aversion to MCMC, rooted mainly in concerns about
sampling time and the presumed difficulty of implementing the required
statistical model \citep{raue_joining_2013, saa_construction_2016}. We
are only aware of two recent attempts to implement a Bayesian kinetic
modelling approach using MCMC. \citet{stapor_pesto_2018} fitted detailed
kinetic models using relatively inefficient MCMC algorithms that do not
scale well to high dimensional parameter spaces limiting the scope of
modelling. Conversely, \citet{st.johnBayesianInferenceMetabolic2018}
utilises an efficient sampling algorithm but uses approximate kinetics,
namely lin-log kinetics \citep{visser_dynamic_2003}, limiting the scope
of interpreting parameters and inferring cellular behaviour in
experimental conditions outside the reference dataset.

There have also been efforts to implement Bayesian inference for kinetic
models without the use of MCMC. Examples of alternative inference
methods include variational inference
\citep{st.johnBayesianInferenceMetabolic2018}, rejection sampling and
approximate Bayesian computation \citep{saa_construction_2016} and
Laplace approximation, in which the Fisher information matrix is used to
calculate a normal approximation around the maximum a posteriori
parameter configuration
\citep{liebermeister_model_2021, gopalakrishnan_k-fit_2020, stapor_pesto_2018, raue_joining_2013}.
Non-MCMC-based Bayesian kinetic models suffer from a lack of reliable
diagnostic tools for verifying that their results approximate the target
posterior distribution. This is a problem because realistic kinetic
models tend to induce highly correlated, non-Gaussian, joint probability
distributions \citep{gutenkunst_2007, stapor_pesto_2018}.

Our application Maud provides the first Bayesian kinetic model to
combine biologically realistic mechanistic accuracy---including accurate
rate laws, post- translational modification and thermodynamics---with
fast, robust MCMC sampling using adaptive Hamiltonian Monte Carlo.
Further, Maud is a general-purpose application that can be used to fit a
wide range of Bayesian kinetic models.

\hypertarget{results-and-discussion}{%
\section{Results and Discussion}\label{results-and-discussion}}

To demonstrate our application's capabilities, we used Maud to analyse
an artificial dataset based on the human methionine cycle. We generated
this dataset using Maud, by simulating measurements for a set of
training and validation conditions based on plausible parameter values.
Next, we performed posterior inference for the training measurements and
prediction for both training and validation measurements.

We investigated Maud's sensitivity to missing measurements by comparing
the results of fitting a full dataset with an intentionally incomplete
dataset. To demonstrate why a full Bayesian approach is preferable to an
approach based on a Laplace approximation of the posterior distribution,
we compared the results of analysing a representative metabolic network
using both methods.

Finally, we dug into our results to find out what our full dataset
methionine model learned about the contributions of different regulatory
factors to the flux through GNMT, an important reaction. This analysis
illustrates how Maud can be used to generate actionable insights about
metabolism without the need for further statistical analysis.

The methionine cycle, illustrated in
Figure~\ref{fig-methionine-reactions}, is a fundamental pathway in human
metabolism, whose intermediate metabolites participate in a variety of
mechanisms which must compete for the same resources. Due to this
competition, as well as the fact that all the functions occur
simultaneously, the methionine cycle is highly regulated, with 6 known
allosteric effectors
\citep{martinov2000, nijhout2006, korendyaseva_allosteric_2008}. This
complex regulation means that quantitative modelling of the methionine
cycle requires a detailed kinetic model: this is why we chose it as a
case study for Maud.

\begin{figure}

\begin{minipage}[t]{\linewidth}

{\centering

\raisebox{-\height}{

\includegraphics{./figures/methionine-reactions.png}

}

}

\end{minipage}%

\caption{\label{fig-methionine-reactions}The methionine cycle as
modelled, with the solid black lines representing the reactions, the
green lines representing allosteric interaction, and the red lines
representing allosteric inhibition. The bold fonts are the reaction
names and the regular font represents the metabolites.}

\end{figure}

\hypertarget{dataset-and-model-specification}{%
\subsection{Dataset and model
specification}\label{dataset-and-model-specification}}

The simulated dataset and underlying kinetic model that we used for our
analysis can be found at
\url{https://github.com/biosustain/Methionine_model/tree/main/data/methionine}
and is described in supporting information section 3.

We constructed a kinetic model of the methionine cycle in Maud's format
using the description in \citet{korendyaseva_allosteric_2008}. The
ordinary differential equation system describing this model is

\begin{align}
\frac{d[met]}{dt} &= v_{Influx} - v_{PROT} - v_{MAT} +v_{MS} + v_{BHMT} \label{eq-meth-ode} \\
\frac{d[amet]}{dt} &= v_{MAT} - v_{GNMT} - v_{METH} \nonumber \\
\frac{d[ahyc]}{dt} &= v_{GNMT} + v_{METH} - v_{AHC} \nonumber \\
\frac{d[hyc-L]}{dt} &= v_{AHC} - v_{CBS} - v_{MS} - v_{BHMT} \nonumber \\
\frac{d[5mthf]}{dt} &= v_{MTHFR} - v_{MS} \nonumber 
\end{align}

After specifying the qualitative aspects of the kinetic model, we
selected parameter values to use as ground truth by Monte Carlo sampling
using a previous model of the methionine cycle as a starting point (see
\citet{saa_construction_2016} for this model).

We used these parameters to simulate steady states in a range of
plausible experimental conditions, again using
\citet{saa_construction_2016} as a starting point. These steady states
were then used to generate simulated measurements using the measurement
model.

For enzyme and metabolite concentration measurements we specified a
standard deviation of 0.1 on natural logarithmic scale, corresponding to
approximately 10\% measurement error. For each reaction measurement a
measurement standard deviation of approximately 10\% of the simulated
value.

These measurement error specifications are somewhat optimistic
considering the many sources of variation and uncertainty affecting
quantitative proteomics, metabolomics and fluxomics analyses, but are a
reasonable first approximation to a realistic set of measurements.

For our main model run, we assumed that all metabolite and enzyme
concentrations were measured, and that there was a reaction measurement
for each of the network's free fluxes.

The simulated experiments and measurements were split into a training
and a validation dataset in a way that achieved a large difference in
flux between the two categories. This was done to evaluate whether the
fitted model is able to extrapolate to conditions well outside the
training dataset rather than merely interpolating between the training
data without necessarily learning the system.

We constructed inputs in Maud's format for each of the analysed
datasets, based on the scenario that the true kinetic model was known
except for parameter values, which needed to be inferred from the
training data and priors. These inputs can be found at
\url{https://github.com/biosustain/Methionine_model/tree/main/data}.

\hypertarget{posterior-inference}{%
\subsection{Posterior inference}\label{posterior-inference}}

The prior distributions and corresponding true parameter values used in
our case study are shown in supporting information section 3.2. They
were chosen to reflect a plausible pre-experimental information state.
In 7 cases, the marginal prior distribution for a parameter disagrees
with the true parameter values used to generate the data; a similar
situation is likely to occur in practice due to \emph{in vivo} vs
\emph{in vitro} measurement differences.

Running standard diagnostic checks indicated that the samples we
generated were from the target posterior distribution. The improved
\(\hat{R}\) statistic
\citep{vehtariRankNormalizationFoldingLocalization2021} for every
variable of interest was within 2\% of 1, indicating appropriate mixing
within and between Markov chains. Additionally, the number of effective
samples was high, indicating that we generated enough posterior samples
to support inferences about the bulks of the distributions of the
sampled parameters. Furthermore, we observed no post warm-up divergent
transitions, indicating that the sampler was able to transform the
log-posterior distribution, avoiding any regions with excessive
curvature that might inhibit exploration via HMC.

Posterior predictive checking indicated that our model achieved a good
fit to the simulated reaction and metabolite concentration measurements,
as shown by the graphs in the top row of Figure~\ref{fig-posterior}.

Analysis of the posterior distributions for kinetic parameters indicated
that these are highly correlated. The marginal posterior distributions
for most kinetic parameters did not shrink significantly compared to the
corresponding marginal prior distributions, even though these
parameters' joint posterior distribution contained enough information to
make accurate out of sample predictions. In some cases, there were
two-dimensional correlations such as the one shown in the bottom left of
Figure~\ref{fig-posterior}; in this case the marginal distribution of
the two parameters is roughly banana shaped. More commonly, however,
two-dimensional pair plots were insufficient to reveal the underlying
correlation structure, as seen in the bottom-right plot in figure
Figure~\ref{fig-posterior}, which shows two marginally independent
parameters. This kind of marginal independence does not mean that the
parameters are independent in the full joint posterior distribution.

These results show that Maud can achieve an adequate fit to a realistic
pathway-sized dataset. This was achieved without fixing the marginal
values of kinetic parameters: the information required to make good
predictions was contained in the correlation structure of the joint
posterior distribution. This finding is consistent with previous
analyses of biological systems that found they are ``sloppy'', that is,
sensitive to parameter combinations rather than marginal parameter
values, with important combinations, scales and regions of sensitivity
being difficult to ascertain in advance
\citep{gutenkunst_2007, poirier_revising_1998}.

The question naturally arises whether the crucial high-dimensional
parameter correlations are linear or non-linear. This is relevant to the
question of model performance, as linear correlations are easier to
correct for. A linearly correlated posterior space would also be easier
to summarise. We address this question in the next section.

\begin{figure}

\begin{minipage}[t]{\linewidth}

{\centering

\raisebox{-\height}{

\includegraphics{./figures/posterior.png}

}

}

\end{minipage}%

\caption{\label{fig-posterior}Marginal posterior distributions from our
main model run. (\textbf{Top left}) Comparison of posterior predictive
intervals with simulated flux measurement values. All the flux
measurements are within the predictive intervals, indicating a good fit.
(\textbf{Top right}) Comparison of posterior predictive intervals with
simulated concentration measurement values. These also show a good fit.
(\textbf{Bottom left}) Pairwise marginal posterior distribution for two
pairwise correlated parameters, namely \(K_m^{AHC1,hcys-L}\) and
\(K_m^{AHC1_ahcys}\). (\textbf{Bottom right}) Pairwise marginal
posterior distribution for two pairwise-uncorrelated parameters, namely
\(K_m^{MAT3,atp}\) and \(K_{cat}^{MAT3}\).}

\end{figure}

\hypertarget{sec-laplace}{%
\subsubsection{Comparison with Laplace
approximation}\label{sec-laplace}}

This current case study illustrates the type of kinetic model and
dataset that Maud can fit. The model we analysed has 10 reactions, 5
state variables and 212 parameters. Generalising from our ability to fit
this model in a reasonable time using Maud, we expect that Maud can be
used to fit realistic Bayesian models of approximately the same size,
but not, for example, genome-scale kinetic models. To fit larger models,
faster steady state solving methods or alternative inference algorithms
will be required. The Laplace approximation, in which the Fisher
information matrix is used to calculate a normal approximation around
the maximum a posteriori parameter configuration, is a popular strategy
for generating approximate posterior samples while avoiding a full MCMC
approach.

Using the recently implemented Laplace approximation in Stan, we were
unable to generate approximate posterior samples for our main methionine
cycle case study using Laplace approximation, as the algorithm could not
recover from solver failures caused by unrealistic parameter
configurations. This is not an inherent issue of the Laplace
transformation but highlights some of the challenges around solving
\eqref{eq-3} for real problems.

In order still to assess the potential use of the Laplace approximation,
we made a comparison for a simpler model. This input can be found at
\url{https://github.com/biosustain/Methionine_model/tree/main/data/example_ode}.
MCMC sampling for this simpler model yielded 800 samples in 625 minutes,
while Laplace sampling yielded the same number of samples in only one
minute. The diagnostics indicated that our algorithm was able to find
the maximum a posteriori parameter configuration, approximate the
Hessian and use these quantities to generate approximate posterior
samples. The results can be found at
\url{https://github.com/biosustain/Methionine_model/tree/main/results}.

Comparing the samples generated using each method shows that the Laplace
approximation does not provide a good approximation to the true
posterior distribution in this case (Figure~\ref{fig-laplace}). As can
be seen from the top left plot, the total log probability density is
clearly different and this was confirmed by a test of the equality of
two empirical univariate distributions (Kolmogorov-Smirnov,
\(p=1.7\times10^{-65}\)).

The difference between the Laplace approximation output and the true
posterior distribution manifests not only in the parameter space, but
also in the measurement space for both training data (upper right) and
validation data (lower left). The lower log likelihood values indication
that the modelled and measured values are further away and that the
Laplace approximation yielded significantly worse predictions than the
true posterior, even for the training data.

To further explore why this is the case we compared samples from the
true posterior and the Laplace approximation for the pairwise marginal
distributions of two Michaelis-Menten constants \(K_{m}^{A,r1}\) and
\(K_{cat}^{r1}\) (Figure~\ref{fig-laplace}, bottom right). This
comparison demonstrates that the Laplace method is not able to capture
the correct relationships between parameters' distributions.

This result shows that MCMC, while slower than Laplace approximation, is
required for posterior inference in this case. We expect that the
Laplace method will produce worse approximations the more complex the
target model. Since the approximation is already unacceptable for our
simple test model, we recommend that Maud users use MCMC sampling in
preference to Laplace approximation if possible when fitting realistic
Bayesian kinetic models.

Our results here also provide circumstantial evidence that the parameter
correlations in Bayesian kinetic model posteriors tend to be non-linear,
as a posterior with only linear correlations would likely be more
germane to Laplace approximation. A conclusion that we drew from this
analysis was that the results of fitting our model cannot be summarised
simply, for example by fitting a multivariate normal distribution to the
posterior draws. We therefore recommend that Maud users store the full
set of MCMC draws rather than using such an approximation. This does not
preclude the possibility that there is an alternative, more compact, way
to summarise the results of Bayesian kinetic model inference; we leave
research into this topic to future work.

\begin{figure}

\begin{minipage}[t]{\linewidth}

{\centering

\raisebox{-\height}{

\includegraphics{./figures/laplace.png}

}

}

\end{minipage}%

\caption{\label{fig-laplace}Graphical comparison of approximate
posterior samples generated using Laplace sampling (blue-green) with
posterior samples generated using MCMC (dark grey). (\textbf{Top left})
the two sets of samples clearly have different marginal distributions
for the overall log probability variable, indicating that the Laplace
samples do not accurately approximate the target distribution.
(\textbf{Top right}) The distribution of marginal posterior predictive
log likelihood values for training data flux measurements shows that the
Laplace method tended to yield much worse predictions compared with the
true model (lower log likelihood values indicate that the modelled and
measured values are further away). (\textbf{Bottom left}) The Laplace
method also tended to produce worse flux predictions for held-back test
measurements. (\textbf{Bottom right}) The marginal joint distribution of
two parameters: \(K_m^{A,rl}\) and \(K_{cat}^{r1}\). The Laplace method
is not able to track the correct joint distribution for this pair of
parameters. This is unsurprising given that the target distribution has
position-dependent scales which are difficult for a linear approximation
to capture.}

\end{figure}

\hypertarget{effect-of-missing-metabolite-concentration-measurements}{%
\subsection{Effect of missing metabolite concentration
measurements}\label{effect-of-missing-metabolite-concentration-measurements}}

To gain insight into our model's robustness to missing measurements, we
also performed a model run with the same 6 experimental datasets, but
with measurements of the metabolite S-Adenosyl-L-homocysteine, or
``ahcys'' removed. Since ahcys regulates three enzymes in the methionine
cycle, including one enzyme which is also thermodynamically regulated,
we expected the removal of these measurements to yield interesting
results.

Comparing model runs with and without the ahcys measurements showed that
Maud can produce sensible results even from incomplete metabolomics
data.

As might be expected, the model with missing measurements did not
correctly infer the missing ahcys concentrations
(Figure~\ref{fig-missing}). Nonetheless, the remaining measured
metabolites were still well predicted, suggesting that information about
the network is still preserved despite the missing measurements.
Comparison of flux measurements in both models also indicated that
removing the ahcys measurement did not result in catastrophic model
failure.

The missing measurements did affect Maud's ability to infer parameter
values correctly (Figure~\ref{fig-missing}, lower left). The model with
full metabolomics learned the true value for the displayed dissociation
constant, despite this value being far from the mean of the
corresponding marginal prior distribution. In contrast, the model with
missing measurements stayed in the neighbourhood of the prior.

This result is reassuring because not having access to all measurements
is a common situation in multi-omics studies. For instance, measuring
all metabolites in a pathway can be infeasible because of limitations of
mass spectrometers, availability of standards, column effects, and
compartmentalisation. However, provided that sufficient information is
available from other sources, our approach can produce sensible results
from incomplete metabolomics data.

\begin{figure}

\begin{minipage}[t]{\linewidth}

{\centering

\raisebox{-\height}{

\includegraphics{./figures/missing.png}

}

}

\end{minipage}%

\caption{\label{fig-missing}Results of removing concentration
measurements for the metabolite \(ahcys_c\) from our case study dataset.
(\textbf{Top}) A comparison of metabolite concentration residuals
between the full measurement dataset (blue-green) and the missing-data
dataset (grey), displayed on natural logarithmic scale. The missing-data
model was unable to estimate the withheld \(ahcys_c\) concentrations.
(\textbf{Bottom Left}) The marginal posterior distribution for the
Michaelis constant \(K_m^{AHC1,ahcys_c}\) in each model, alongside the
true parameter value used to generate both datasets. The true value is
recovered by the complete-data model but not by the missing-data model.
(\textbf{Bottom Right}) The distribution of total log-likelihood for
out-of-sample flux measurements in both models. There is a significant
overlap between the two distributions, suggesting that removing the
\(ahcys_c\) measurement did not cause catastrophic prediction failure.
However, overall the complete-data model tended to make better
predictions.}

\end{figure}

\hypertarget{application-to-regulatory-understanding}{%
\subsection{Application to regulatory
understanding}\label{application-to-regulatory-understanding}}

To demonstrate how Maud's output can be used to yield useful metabolic
insights we used the results of our case study to explain why the flux
of the enzyme GNMT is higher in dataset 1 than in dataset 12
(Figure~\ref{fig-decomposition}). GNMT is an irreversible enzyme that is
homotropically activated by its substrate, competitively inhibited by
its product and heterotropically inhibited by 5,10-
methylenetetrahydrofolate (mlthf). The complex regulation makes it the
ideal test case to elucidate regulatory changes.

Regulation can be separated into enzyme abundance, allostery and
saturation, and we can plot the marginal posterior distribution of the
ratio of the corresponding regulatory component in dataset 1 compared
with dataset 12 (Figure~\ref{fig-decomposition}, right panel). A
positive value indicates that the component was increased in dataset 1
relative to dataset 12, with 0 indicating no difference. The
probability, according to our model, of the component acting in each
direction is given by the relative area under the curve on each side of
the zero point.

Our model correctly inferred that saturation and allosteric effects were
the main drivers of regulation between the two datasets in this case,
with the curves for each component aligning with the ground truth shown
in red.

Importantly, this form of analysis takes into account all modelled
sources of uncertainty, including uncertainty about the measured values
of the flux in each dataset. Our result shows that Maud could be used in
this realistic case not only to explain an observed difference in
fluxes, but also to provide a reasonable judgement as to the
explanation's robustness.

\begin{figure}

\begin{minipage}[t]{\linewidth}

{\centering

\raisebox{-\height}{

\includegraphics{./figures/decomposition.png}

}

}

\end{minipage}%

\caption{\label{fig-decomposition}Illustration of how analysing a system
with Maud can yield actionable insights about the underlying metabolic
network. (\textbf{Top Left}) A schematic of the regulatory interactions
associated with the enzyme \(GNMT1\). Dashed green lines represents
allosteric activation, dashed red lines indicate allosteric inhibition
and solid red lines represent competitive inhibition. (\textbf{Bottom
Left}) Comparison of marginal posterior distributions for \(GNMT\) flux
in datasets 1 and 2. (\textbf{Right}) Log-scale ratios of the regulatory
elements defined in \eqref{eq-decomposition}. Note that the
\(Reversibility\) and \(k_cat\) components are excluded: this is because
this reaction was modelled as irreversible, and \(k_{cat}\) was modelled
as constant across datasets. These plots identify why flux in dataset 12
is higher than in dataset 1: the flux increase is due to allostery and
saturation with no control from enzyme concentration changes.}

\end{figure}

\hypertarget{methods}{%
\section{Methods}\label{methods}}

Maud is a command line application implementing Bayesian inference for a
wide range of realistic kinetic models. Maud is written in Python
\citep{vanrossumPythonReferenceManual2009}, designed for use on Windows,
macOS and Linux, is \texttt{pip} installable from the Python Package
Index as \texttt{maud-metabolic-models}, documented at
\url{https://maud-metabolic-models.readthedocs.io} and actively
developed and maintained at \url{https://github.com/biosustain/Maud/}.

To use Maud, a user must first collate appropriate input information,
represent it in files with Maud's required formats (see supporting
information section on input format). Maud's command line interface
provides commands for inference via a range of algorithms including
adaptive Hamiltonian Monte Carlo, as well as commands for simulation and
making out-of-sample predictions. Results are stored in files, using a
structured, interoperable format.

As well as parameter values, Maud also performs inference for derived
quantities including the components of the regulatory decomposition
described below in \eqref{eq-decomposition}, log likelihoods, simulated
measurement values and metabolic control analysis coefficients
\citep{kacser_control_1973}.

In the rest of this section we describe Maud's kinetic model at high
level and discuss Maud's statistical model, implementation and how it
solves the crucial steady state problem. Further details about Maud's
kinetic model can be found in the supporting information.

\hypertarget{kinetic-model}{%
\subsection{Kinetic model}\label{kinetic-model}}

Maud's kinetic model decomposes into factors:

\begin{equation}
F(C;\theta) = Enzyme\cdot k_{cat}\cdot Reversibility \cdot Saturation \cdot Allostery \label{eq-decomposition}
\end{equation}

Each of the terms on the right-hand side of \eqref{eq-decomposition} is
a function of \(C\) and \(\theta\). This idea is taken from
\citet{noor_note_2013}. The terms usefully gather physically meaningful
and conceptually distinct factors contributing to reaction fluxes.
\(Enzyme\) captures the effect of enzyme concentration, \(k_{cat}\) that
of enzyme efficiency, \(Reversibility\) quantifies thermodynamic
effects, \(Saturation\) the effect of enzyme availability and
\(Allostery\) the effect of post translational modifications.

We used the model of enzyme saturation from
\citet{liebermeister_modular_2010} and the generalised
Monod-Wyman-Changeux model of Allosteric regulation introduced in
\citep{monod_nature_1965, changeux_2013, popova_generalization_1975, popova_description_1979}
and used more recently in \citet{matosGRASPComputationalPlatform2022}.
To capture the effect on enzyme activity of coupled phosphorylation and
dephosphorylation processes we developed a new mathematical model
inspired by the generalised MWC model of allosteric regulation. Full
details of all mathematical aspects of Maud's kinetic model can be found
in supporting information section 2.

\hypertarget{statistical-model}{%
\subsection{Statistical model}\label{statistical-model}}

Maud represents information from measurements using generalised linear
regression models that probabilistically connect realised measurements
with true values of the measurable quantities. Information from other
sources is represented using a prior distribution over a set of latent
parameters. The parameters and the measurable quantities are connected
by a generative model encompassing Maud's kinetic model as well as the
steady state equation \eqref{eq-3}. Together, the measurement model,
prior model and generative model determine a joint probability function
that assigns a probability density to any possible combination of
measurements and parameters.

Below we describe the prior and measurement models, as the generative
model has already been discussed above.

\hypertarget{prior-model}{%
\subsubsection{Prior model}\label{prior-model}}

Maud's prior model includes unknown parameters corresponding to
quantities in the kinetic model that are assumed to be unknown, other
than steady state metabolite concentrations and fluxes, which are
derived from the values of other parameters by solving the steady state
problem. See table 1 in this paper's supporting information for a
description of all these parameters and their dimensions. Note that some
quantities in Maud's kinetic model are not treated as parameters: for
example, temperatures, compartment volumes and the formation energy of
water. Maud treats these quantities as if they were known precisely:
they can be configured by the user or default values can be used.
Although in practice there can be considerable uncertainty regarding
these quantities, we chose to disregard this uncertainty in the interest
of simplicity.

Except for metabolites' standard condition Gibbs energy changes of
formation, Maud uses independent normal prior distributions for
parameters that can in principle be both negative and positive. For
parameters that are constrained to be positive, Maud uses independent
log-normal distributions. Formation energy parameters have a
multivariate normal prior distribution. Location, scale and covariance
parameters for all these prior distributions can be selected freely by
the user.

\hypertarget{measurement-model}{%
\subsubsection{Measurement model}\label{measurement-model}}

Maud's measurement model considers three types of measurement:
metabolite concentration measurements, enzyme concentration measurements
and flux measurements, represented by vectors \(𝑦^{𝑐𝑜𝑛𝑐}\) \(𝑦^{𝑒𝑛𝑧}\)
and \(𝑦^{𝑓𝑙𝑢𝑥}\) respectively.

All measurements are specific to an experimental condition; that is, a
case where the true state of the network, including knockouts, boundary
conditions and state variables as well as kinetic and thermodynamic
parameters, can safely be assumed to be the same. Maud's statistical
model allows for arbitrarily many experimental conditions, and for any
measurable quantity to be measured any number of times in any condition.

Metabolite and enzyme measurements are intended to represent the results
of quantitative metabolomics and proteomics experiments. The likelihood
functions for such measurements are:

\begin{align}
y_i^{conc} &\sim LN(\ln{\hat{y}_i^{conc}}, \sigma_i^{conc})\label{eq-yconc} \\
y_i^{enz} &\sim LN(\ln{\hat{y}_i^{enz}}, \sigma_i^{enz})\label{eq-yenz}
\end{align}

Both equations are log-normal generalised linear models with a standard
link function (the natural logarithm \(\ln\)) and known standard
deviation \$ \sigma\_\{𝑐𝑜𝑛𝑐\}\$. The use of this measurement model is
motivated by the consideration that concentrations are constrained to be
non-negative, so the measurement model should avoid assigning positive
probability mass to negative metabolite concentration values. In
addition, we expect the precision of most metabolomics and proteomics
experiments to be roughly proportional to the true value of the measured
quantity, which supports a measurement model with constant coefficient
of variation. The measurement standard deviations \(\sigma_{𝑐𝑜𝑛𝑐}\) and
\(\sigma_{𝑒𝑛𝑧}\) are assumed to be known exactly for simplicity;
plausible values can be elicited by considering the likely coefficient
of variation of the measuring apparatus. Our measurement model improves
on analyses of metabolomics and proteomics data that assume a regression
model with normally distributed errors, whether explicitly using a
standard linear model or implicitly using ordinary least squares
fitting.

Flux measurements, representing the results of quantitative fluxomics
analyses, are modelled using a likelihood function from a standard
linear regression model \eqref{eq-yflux}. Flux measurements can be
obtained from isotope labelling experiments using metabolic flux
analysis, for example as described in (Young 2014). When entering flux
measurements, it is important only to specify measurements for a
network's free fluxes, as the values of some steady state fluxes in a
metabolic network are constrained by others, with the result that
dependent fluxes cannot typically be measured separately. If
measurements of multiple dependent fluxes are entered, information will
inappropriately be double counted.

\begin{equation}
y_i^{flux} \sim LN(\ln{\hat{y}_i^{flux}}, \sigma_i^{flux})\label{eq-yflux}
\end{equation}

The use of independent measurement models for metabolite, enzyme and
flux measurements carries an implicit assumption that there are no
systematic correlations in the measurement errors. This choice was
motivated by simplicity - it would be better to use a model with
potentially correlated measurements. Similarly, it would be preferable
to include measurement errors as model parameters, thereby avoiding
possible bias due to incorrect assessments of measurement accuracy.
However, we chose to use a simpler measurement model to avoid the
complexity and potential fitting issues that these changes would entail.

Finally, the reader may wonder why Maud uses a linear regression model
for reaction flux measurements even though this creates the potential
for erroneous double counting and requires non-trivial upstream
modelling, as intracellular fluxes typically cannot be measured
directly. Instead, fluxes are typically inferred from isotope labelling
experiments using metabolic flux analysis: see
\citet{daiUnderstandingMetabolismFlux2017} for more about this method.
Ideally Maud's measurement model for fluxes would extend from fluxes to
the results of potential labelling experiments, thereby removing the
need for upstream analysis and avoiding any double counting. This option
has not yet been pursued, again for the sake of simplicity.

\hypertarget{implementation}{%
\subsection{Implementation}\label{implementation}}

Maud uses the Python library click
\citep{clickdevelopersClickPythonComposable2022} to implement a command
line interface. The command line interface loads input files as Python
dictionaries, which are parsed using the Python library \texttt{toml}
\citep{pearsonTomlPythonLibrary2020} and then validated and converted
into structured \texttt{MaudInput} objects using Pydantic
\citep{pydanticdevelopersPydantic2022}. Maud's statistical model is
implemented in the probabilistic programming language Stan
\citep{carpenterStanProbabilisticProgramming2017} and accessed using the
interface cmdstanpy \citep{standevelopmentteamCmdStanPy2022}. For
posterior sampling, Maud uses the \texttt{MaudInput} to create an input
file for Stan and then trigger posterior sampling using adaptive
Hamiltonian Monte Carlo. See
\citet{betancourtConceptualIntroductionHamiltonian2018} for more about
this algorithm.

When sampling is complete, Maud converts to the output into the standard
format \texttt{InferenceData} using the Python library arviz
\citep{kumarArviZUnifiedLibrary2019} and saves it as a json file, along
with some information for debugging. This format allows for easy
checking of MCMC diagnostics including divergent transitions, effective
sample size and the improved \(\hat{R}\) statistic proposed in
\citet{vehtariRankNormalizationFoldingLocalization2021}.

\hypertarget{solving-the-steady-state-problem}{%
\subsection{Solving the steady state
problem}\label{solving-the-steady-state-problem}}

Maud's parameters are connected with measurable quantities via the
steady state equation \eqref{eq-3}. Posterior sampling using adaptive
Hamiltonian Monte Carlo requires repeatedly solving \eqref{eq-3} and
finding the gradients of its solution with respect to all model
parameters. This must be done numerically as analytic solutions are not
available for realistic kinetic models.

The speed at which this problem can be solved is tightly coupled with
the size and complexity of metabolic network that can practically be
modelled. See \citet{timonenImportanceSamplingApproach2022a} for more
about considerations involved in this kind of modelling.

To solve \eqref{eq-3} and find its gradients, Maud uses a hybrid method
involving two numerical solvers from the SUNDIALS suite
\citep{serbanCVODESSensitivityEnabledODE2005}: CVODES and IDAS via their
interface from Stan. The hybrid method follows that proposed by
\citet{margossianComputingSteadyStates2018}, and involves numerically
evolving the ODE system for a short period of time, then using the
difference between the evolved and starting concentrations as the target
for a numerical algebra solver.

\hypertarget{supporting-information}{%
\section{Supporting information}\label{supporting-information}}

The analyses described in this paper, as well as instructions for
reproducing our results, figures and paper, can be found at
\url{https://biosustain/Methionine_model}.

The document \texttt{supporting\_information.pdf} contains the following
information:

\begin{itemize}
\tightlist
\item
  Description of Maud's input format.
\item
  Description of Maud's kinetic model including rate equations as well
  as parameters and their dimensions.
\item
  Details of the procedure used to generate our case study results,
  including generating the artificial case study datasets, specifying
  priors and carrying out computation.
\item
  Specification of the prior distribution used for the case studies.
\end{itemize}

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

This research was funded by the Novo Nordisk Foundation (Grant numbers
NNF20CC0035580 NNF14OC0009473).

\hypertarget{references}{%
\section{References}\label{references}}

\renewcommand{\bibsection}{}
\bibliography{bibliography.bib}

\end{document}