### 2024-02-12 paper/supporting_information.qmd
Final resubmission changes

---
affiliations:
  - id: biosustain
    name: "DTU"
    department: "The Novo Nordisk Foundation Center for Biosustainability"
    city: "Kongens Lyngby 2800"
    country: "Denmark"
  - id: uq
    name: "The University of Queensland"
    department: "Australian Institute for Bioengineering and Nanotechnology (AIBN)"
    city: "St Lucia 4067"
    country: "Australia"
author: 
  - name: "Teddy Groves"
    orcid: "0000-0002-7109-3270"
    email: "tedgro@biosustain.dtu.dk"
    affiliations:
      - ref: biosustain
  - name: "Nicholas Luke Cowie"
    affiliations:
      - ref: biosustain
    orcid: "0000-0003-3367-2105"
  - name: "Lars Keld Nielsen"
    affiliations:
      - ref: uq
      - ref: biosustain
    orcid: "0000-0001-8191-3511"
title: Supporting information - Bayesian regression facilitates quantitative modelling of cell metabolism
format:
  html:
    embed-resources: true
    include-in-header: mathjax.html
  acs-pdf: 
    include-in-header: 
      - text: \renewcommand{\thepage}{S\arabic{page}}
    keep-tex: true
    journal:
      id: asbcd6  #This is an identifier for the target journal: from https://pubs.acs.org/, the identifier is the part of the URL after https://pubs.acs.org/journal/, e.g. for J. Org. Chem. it would be joceah.
      type: article  #article, note, communication, review, letter, or perspectivee
      layout: traditional #traditional or twocolumn, twocolumn currently broken for longtables
  docx: default
number-sections: true
bibliography: bibliography.bib
---

{{< pagebreak >}}

This document provides information in support of our article "Bayesian
regression facilitates quantitative modelling of cell metabolism".

The results of all reported Maud runs can be found at <https://github.com/biosustain/Methionine_model/blob/main/results>.

# Maud's input format

Maud inputs are structured directories, somewhat inspired by the PEtab
format [@SchmiesterSch2021]. A Maud input directory must contain a toml
[@preston-wernertomandgedampradyunTOMLSpecification0rc2020] file called
`config.toml` which gives the input a name, configures how Maud will be run
and tells Maud where to find the other files, allowing these to have custom
names. It must also include a file containing a kinetic model definition,
a file specifying information about parameters and a file with information
experiments. The required structure of these files is documented at <https://maud-metabolic-models.readthedocs.io/en/latest/inputting.html>.
The input is validated against a Pydantic [@pydanticdevelopersPydantic2022] 
data model.

We chose to implement a custom input format despite the existence of standard
formats in similar areas, including SBML [@keatingSBMLLevelExtensible2020]
and PEtab [@SchmiesterSch2021]. This choice was partly motivated by the need
to ensure flexibility as Maud was developed, but there are also features of
SBML and PEtab that make them structurally unsuitable in this context. Our
requirements for an input format included that it be mathematics-free, so that
all mathematical details are encapsulated in source code, and that it has a
detailed, verifiable structure. These requirements made toml more attractive
than SBML: toml is easier for humans to read and edit and can straightforwardly
be validated using tools like Pydantic. Further, an SBML representation of our
desired input would not contain differential equations.  It would therefore not
be interoperable with most SBML targeting software, which typically assumes that
differential equations are available and does not know about Maud’s structure.

# Maud's kinetic model

## Parameters

@tbl-params shows all of Maud's unknown parameters along with their dimensions

Note that Maud’s metabolic model includes some quantities that are not treated
as parameters in its statistical model, including temperatures, compartment
volumes and the formation energy of water. Maud treats these quantities as
if they were known precisely: they can be configured by the user or default
values can be used. Although in practice there can be considerable uncertainty
regarding these quantities, we chose to disregard this uncertainty in the
interest of simplicity.

| Parameter | Modelled quantity | Dimensions | 
| --------- | ----------------- | ---------- |
| $\Delta_fG$| Formation energy | metabolites |
| $k_M$ | Michaelis Menten constants | Substrates of all enzyme/reactions and products of reversible enzyme/reactions |
| $k_I$ | Inhibition constants | Inhibiting metabolite/compartments of enzyme/reactions exhibiting competitive inhibition |
| $k_{cat}$ | Rate constants | Enzyme/reactions |
| $L_0$ | Transfer constants | Allosteric interactions |
| $e_T$ | T dissociation constants | Modifying metabolites of allosteric inhibitions |
| $e_R$ | R dissociation constants | Modifying metabolites of allosteric activations |
| $k_{cat\, pme}$ | Rate constants of phosphorylation modifying enzymes | Phosphorylation modifying enzymes |
| $v_{drain}$ | Drain fluxes | Drains, experiments |
| $Enzyme$ | Enzyme concentrations | Enzymes, experiments |
| $C_{unbalanced}$ | Unbalanced metabolite/compartment concentrations | Unbalanced metabolite/compartments, experiments |
| $C_{pme}$ | Phosphorylation modifying enzyme concentrations | Phosphorylation modifying enzymes, experiments|
| $\psi$ | Membrane potentials | Experiments |

: Table S1 – Parameters of Maud's statistical model {#tbl-params}

Solving the steady state problem for a given set of parameters in an experiment
yields a vector $C_{balanced}$ of balanced metabolite concentrations. These
are combined with the balanced metabolite concentrations $C_{unbalanced}$ to
produce a vector $C_{mic}$ with a concentration for each metabolite/compartment
combination.

$\Delta_fG$ parameters can optionally be fixed; this can be useful for
computational purposes, as for example to avoid estimating the formation energy
of a metabolite about which there is no available information due to it only
participating in irreversible reactions.

## Rate equations

As discussed in the main text, Maud’s kinetic model decomposes into factors
contributing to the flux in a metabolic network in an experiment as shown in
equation \eqref{eq-decomposition}. For succinctness, and since Maud's model
assumes that there are no interactions between experiments, we omit any notation
referring to experiments below. We also omit any reference to the network's
drain reactions: these are modelled as being exactly determined by the values of
the parameter vector $v_{drain}$.

\begin{equation}
F(C;\theta) = Enzyme\cdot k_{cat}\cdot Reversibility \cdot Saturation \cdot Allostery \label{eq-decomposition}
\end{equation}

The term $Enzyme$ in equation \eqref{eq-decomposition} is a vector of
non-negative real numbers representing the concentration of the enzyme
catalysing each reaction.

The term $k_{cat}$ in equation \eqref{eq-decomposition} is a vector of
non-negative real numbers representing the  amount of flux carried per unit of
saturated enzyme.

The term $Reversibility$ in equation \eqref{eq-decomposition} is a vector of
real numbers capturing the impact of thermodynamic effects on the reaction's
flux, as shown in equation \eqref{eq-reversibility}.

\begin{align} 
  Reversibility &= 1 - \exp(\frac{\Delta_{r}G + RT \cdot S^T \ln(C_{mic})}{RT}) \label{eq-reversibility} \\ 
  \Delta_{r}G &= S^{T}\Delta_{f}G + n F \psi \nonumber 
\end{align}

The terms in \eqref{eq-reversibility} have the following meanings:

- $T$ is the temperature in Kelvin (a number), 
- $R$ is the gas constant (a number), 
- $\Delta_rG$ is a vector representing the Gibbs free energy change of each reaction in standard conditions, 
- $\Delta_fG$ is a vector representing the standard condition Gibbs free energy change of each metabolite's formation reaction, or in other words each metabolite's 'formation energy'. 
- $n$ is a vector representing the number of charges transported by each reaction. 
- $F$ is the Faraday constant (a number) 
- $\psi$ is a vector representing each reaction's membrane potential (these numbers only matter for reactions that transport non-zero charge)

Note that, for reactions with zero transported charge, the thermodynamic effect
on each reaction is derived from metabolite formation energies. This formulation
is helpful because, provided that all reactions' rates are calculated from the
same formation energies, they are guaranteed to be thermodynamically consistent.

The term $n$ accounts for both the charge and the directionality. For instance,
a reaction that exports 2 protons to the extracellular space in the forward
direction would have -2 charge. If a negatively charged molecule like acetate is
exported in the forward direction, $n$ would be 1.

Note that this way of modelling the effect of transported charge does not take
into account that the concentration gradient used by the transport is that of
the dissociated molecules. Thus, this expression is only correct for ions whose
concentration can be expressed in the model only in the charged form; e.g.,
protons, $K^+$, $Na^+$, $Cl^-$, etc.

The term $Saturation$ in equation \eqref{eq-decomposition} is a vector of
non-negative real numbers representing, for each reaction, the fraction of
enzyme that is saturated, i.e. bound to one of the reaction's substrates.
To describe saturation we use equation \eqref{eq-saturation}, which is taken
from @liebermeister_modular_2010 and @noor_note_2013. Additionally, this term captures competitive
inhibition: as competitive inhibitor concentration increases, the saturation
denominator increases, effectively decreasing the saturation of the substrate on
the total enzyme pool. Conversely, as the substrate concentration increases this
term approaches 1.

\begin{align} 
Saturation_r &= a \cdot \text{free enzyme ratio}\label{eq-saturation} \\ 
a &= \prod_{\text{s substrate}}\frac{C_{mic}^s}{k_{M}^{rs}} \nonumber \\ 
\text{free enzyme ratio} &= \begin{cases} \begin{aligned}
  &\prod_{\text{s sustrate}} (1 + \frac{C_{mic}^s}{k_{M}^{rs}})^{S_sr} \\
  + &\sum_{\text{c inhibitor}}\frac{C_{mic}^c}{k_I^{rc}}\end{aligned} & r\text{ irreversible} \\ 
  \begin{aligned}
  -1  &+ \prod_{\text{s sustrate}} (1 + \frac{C_{mic}^s}{k_{M}^{rs}})^{S_sr} \\
      &+ \prod_{\text{p product}} (1 + \frac{C_{mic}^p}{k_{M}^{rp}})^{S_pr} \\ 
      &+ \sum_{\text{c inhibitor}}\frac{C_{mic}^c}{k_I^{rc}} 
  \end{aligned}
    & r\text{ reversible} 
    \end{cases} \nonumber
\end{align}

The term $Allostery$ in equation \eqref{eq-decomposition} is a vector of
non-negative numbers describing the effect of allosteric regulation on each
reaction. Allosteric regulation happens when binding to a certain molecule
changes an enzyme's shape in a way that changes its catalytic behaviour. We
use equation \eqref{eq-allostery} to describe this phenomenon, following the
generalised MWC approach described in @monod_nature_1965, @changeux_2013,
@popova_generalization_1975 and @popova_description_1979.

\begin{align} 
  Allostery_r &= \frac{1}{1 + L_0^r \cdot (\text{free enzyme ratio}_r \cdot \frac{Qtense}{Qrelaxed})^{subunits}}\label{eq-allostery} \\ 
       Qtense &= 1 + \sum_{\text{i inhibitor}} \frac{C_{mic}^i}{e_T^{ri}}\nonumber \\ 
     Qrelaxed &= 1 + \sum_{\text{a activator}} \frac{C_{mic}^a}{e_R^{ra}}\nonumber
\end{align}

The parameter $L_0$ in equation \eqref{eq-decomposition} is called the transfer
constant, and the parameter vectors $e_T$ and $e_R$ are called tense and relaxed
dissociation constants respectively.

Finally, the term $Phosphorylation$ in equation \eqref{eq-decomposition}
captures the important effect whereby enzyme activity is altered due to a
coupled process of phosphorylation and dephosphorylation. This description
achieves a similar behaviour to the MWC formalism for describing allosteric
regulation, but using the rates of phosphorylation and dephosphorylation rather
than concentrations of metabolites.

\begin{align}
Phosphorylation_r &= (\frac{\alpha}{\alpha + \beta})^{subunits} \label{eq-phosphorylation}\\
\alpha &= \sum_{\text{p phosphoylator}} k_{cat\,pme}^{p} \cdot C_{pme}^p \nonumber \\
\beta &= \sum_{\text{d dephosphoylator}} k_{cat\,pme}^{d} \cdot C_{pme}^d \nonumber 
\end{align}

# Methionine case study {#sec-methionine-case-study}

## Dataset generation

Starting with the model in @saa_construction_2016, we extracted values for
enzyme concentrations, boundary conditions and fluxes. We used these values
to generate MCMC samples using Maud using the priors specified in section
@sec-methionine-priors. When this was finished, we selected one sample with
relatively high log probability to use as a ground truth in our case study.
These parameter values are shown below in table @tbl-case-study-params. We
manually inspected the parameter values to screen for any obviously implausible
values; we did not find any of these.

## Prior distributions compared with true parameter values {#sec-methionine-priors}

@tbl-case-study-params shows the prior distributions we used for independent
parameters. The first two columns show the 1% and 99% quantiles of each
marginal prior distribution. True parameter value are shown in column three,
and the last column shows the z-score on log scale of the true parameter value
according the marginal prior distribution. As can be seen from the table, there
are 7 parameters for which the true value is outside the 1%-99% range.

| parameter name                      | 1% prior quantile | 99% prior quantile | true value   | prior Z-score of true value |
| ----------------------------------- | ----------------- | ------------------ | ------------ | --------------------------- |
| $e_{𝑅}^{𝐶𝐵𝑆1,𝑎𝑚𝑒𝑡𝑐}$              | 3.430e-06      | 0.002480           | 9.3e-05 | 0.004                       |
| $e_{𝑅}^{𝐺𝑁𝑀𝑇1,𝑎𝑚𝑒𝑡𝑐}$            | 3.000e-05      | 0.002000           | 2.000e-05 | -2.787                     |
| $e_{𝑅}^{𝑀𝐴𝑇3,𝑎𝑚𝑒𝑡𝑐}$              | 1.000e-04      | 0.001000           | 3.170e-04 | 0.003                       |
| $e_{𝑅}^{𝑀𝐴𝑇3,𝑚𝑒𝑡−𝐿𝑐}$             | 4.500e-04      | 0.000800           | 6.000e-04 | 0.000                       |
| $e_{𝑅}^{𝑀𝑇𝐻𝐹𝑅1,𝑎ℎ𝑐𝑦𝑠𝑐}$         | 1.120e-07      | 0.000081           | 2.000e-06 | -0.101                     |
| $e_{𝑇}^{𝐺𝑁𝑀𝑇1,𝑚𝑙𝑡ℎ𝑓𝑐}$           | 1.120e-05      | 0.008050           | 2.290e-04 | -0.136                     |
| $e_{𝑇}^{𝑀𝑇𝐻𝐹𝑅1,𝑎𝑚𝑒𝑡𝑐}$          | 1.120e-07      | 0.000081           | 1.500e-05 | 0.549306                    |
| $k_{𝑐𝑎𝑡}^{𝐴𝐻𝐶1}$                     | 1.200e+02      | 400.000000         | 2.340e+02 | 0.179861                    |
| $k_{𝑐𝑎𝑡}^{𝐵𝐻𝑀𝑇1}$                   | 6.000e+00      | 35.000000          | 1.380e+01 | -0.135                     |
| $k_{𝑐𝑎𝑡}^{𝐶𝐵𝑆1}$                     | 1.000e+01      | 188.000000         | 7.020e+00 | -2.887                     |
| $k_{𝑐𝑎𝑡}^{𝐺𝑁𝑀𝑇1}$                   | 7.000e-01      | 60.000000          | 1.050e+01 | 0.352083                    |
| $k_{𝑐𝑎𝑡}^{𝑀𝐴𝑇1}$                     | 8.200e-02      | 59.100000          | 7.900e+00 | 0.44375                     |
| $k_{𝑐𝑎𝑡}^{𝑀𝐴𝑇3}$             | 5.890e-01      | 424.000000         | 1.990e+01 | 0.080556                    |
| $k_{𝑐𝑎𝑡}^{𝑀𝐸𝑇𝐻−𝐺𝑒𝑛}$ | 4.840e-01      | 349.000000         | 1.160e+00 | -1.209                     |
| $k_{𝑐𝑎𝑡}^{𝑀𝑆1kcatMS1}$                | 1.000e+00      | 3.300000           | 1.770e+00 | -0.091                     |
| $k_{𝑐𝑎𝑡}^{𝑀𝑇𝐻𝐹𝑅1}$       | 1.300e+00      | 4.200000           | 3.170e+00 | 0.183333                    |
| $k_{𝑐𝑎𝑡}^{𝑃𝑅𝑂𝑇1}$                   | 1.590e-01      | 0.222000           | 2.650e-01 | 0.41875                     |
| $k_{𝐼}^{𝐺𝑁𝑀𝑇1,𝑎ℎ𝑐𝑦𝑠𝑐}$           | 2.000e-06      | 0.001400           | 5.300e-05 | 0.010                       |
| $k_{𝐼}^{𝑀𝐴𝑇1,𝑎𝑚𝑒𝑡𝑐}$              | 3.000e-04      | 0.000400           | 3.470e-04 | 0.014                       |
| $k_{𝐼}^{𝑀𝐸𝑇𝐻−𝐺𝑒𝑛,𝑎ℎ𝑐𝑦𝑠𝑐}$     | 1.000e-06      | 0.000030           | 6.000e-06 | 0.021                       |
| $k_{𝑀}^{𝐴𝐻𝐶1,𝑎ℎ𝑐𝑦𝑠𝑐}$             | 5.220e-05      | 0.037600           | 2.320e-05 | -2.050                     |
| $k_{𝑀}^{𝐴𝐻𝐶1,𝑎𝑑𝑛𝑐}$                | 1.670e-07      | 0.000120           | 5.660e-06 | 0.081944                    |
| $k_{𝑀}^{𝐴𝐻𝐶1,ℎ𝑐𝑦𝑠−𝐿𝑐}$            | 1.580e-07      | 0.000114           | 1.060e-05 | 0.318056                    |
| $k_{𝑀}^{𝐵𝐻𝑀𝑇1,ℎ𝑐𝑦𝑠−𝐿𝑐}$          | 1.200e-05      | 0.000032           | 1.980e-05 | 0.049                       |
| $k_{𝑀}^{𝐵𝐻𝑀𝑇1,𝑔𝑙𝑦𝑏𝑐}$            | 4.720e-05      | 0.034000           | 8.460e-03 | 0.659028                    |
| $k_{𝑀}^{𝐶𝐵𝑆1,ℎ𝑐𝑦𝑠−𝐿𝑐}$            | 1.000e-06      | 0.000025           | 4.240e-05 | 3.090                       |
| $k_{𝑀}^{𝐶𝐵𝑆1,𝑠𝑒𝑟−𝐿𝑐}$             | 2.000e-06      | 0.000004           | 2.830e-06 | 0.004                       |
| $k_{𝑀}^{𝐺𝑁𝑀𝑇1,𝑎𝑚𝑒𝑡𝑐}$            | 1.300e-05      | 0.009400           | 5.200e-04 | 0.1375                      |
| $k_{𝑀}^{𝐺𝑁𝑀𝑇1,𝑎ℎ𝑐𝑦𝑠𝑐}$           | 4.100e-07      | 0.000295           | 1.100e-05 | 0.000                       |
| $k_{𝑀}^{𝐺𝑁𝑀𝑇1,𝑔𝑙𝑦𝑐}$              | 5.480e-05      | 0.039500           | 2.540e-03 | 0.189583                    |
| $k_{𝑀}^{𝐺𝑁𝑀𝑇1,𝑠𝑎𝑟𝑐𝑠𝑐}$          | 3.730e-09      | 0.000003           | 1.000e-07 | 0.000                       |
| $k_{𝑀}^{𝑀𝐴𝑇1,𝑚𝑒𝑡−𝐿𝑐}$             | 1.400e-05      | 0.000720           | 1.070e-04 | 0.074                       |
| $k_{𝑀}^{𝑀𝐴𝑇1,𝑎𝑡𝑝𝑐}$                | 5.270e-05      | 0.038000           | 2.030e-03 | 0.125694                    |
| $k_{𝑀}^{𝑀𝐴𝑇3,𝑚𝑒𝑡−𝐿𝑐}$             | 4.470e-05      | 0.032200           | 1.130e-03 | -0.029                     |
| $k_{𝑀}^{𝑀𝐴𝑇3,𝑎𝑡𝑝𝑐}$                | 5.270e-05      | 0.038000           | 2.370e-03 | 0.179167                    |
| $k_{𝑀}^{𝑀𝐸𝑇𝐻−𝐺𝑒𝑛,𝑎𝑚𝑒𝑡𝑐}$      | 7.000e-06      | 0.000013           | 9.370e-06 | -0.135                     |
| $k_{𝑀}^{𝑀𝑆1,5𝑚𝑡ℎ𝑓𝑐}$                | 3.320e-06      | 0.002390           | 6.940e-05 | -0.124                     |
| $k_{𝑀}^{𝑀𝑆1,ℎ𝑐𝑦𝑠−𝐿𝑐}$              | 1.000e-06      | 0.000003           | 1.710e-06 | -0.054                     |
| $k_{𝑀}^{𝑀𝑇𝐻𝐹𝑅1,𝑚𝑙𝑡ℎ𝑓𝑐}$         | 7.500e-05      | 0.000088           | 8.080e-05 | -0.158                     |
| $k_{𝑀}^{𝑀𝑇𝐻𝐹𝑅1,𝑛𝑎𝑑𝑝ℎ𝑐}$         | 1.600e-05      | 0.000028           | 2.090e-05 | -0.105                     |
| $k_{𝑀}^{𝑃𝑅𝑂𝑇1,𝑚𝑒𝑡−𝐿𝑐}$           | 4.500e-05      | 0.000085           | 4.390e-05 | -2.507                     |
| $𝐿_{0}^{𝐶𝐵𝑆1}$                          | 3.730e-02      | 26.800000          | 1.030e+00 | 0.017                       |
| $𝐿_{0}^{𝐺𝑁𝑀𝑇1}$                        | 3.730e-02      | 26.800000          | 1.310e+02 | 0.3875                      |
| $𝐿_{0}^{𝑀𝐴𝑇3}$                          | 3.730e-03      | 2.680000           | 1.080e-01 | 0.037                       |
| $𝐿_{0}^{𝑀𝑇𝐻𝐹𝑅1}$                      | 1.120e-01      | 80.500000          | 3.920e-01 | -1.018                     |

: Table S2 – Parameter specification, marginal prior distributions and true parameter values used in our case study. {#tbl-case-study-params}

$\Delta_fG$ parameters for most metabolites were fixed; those that were
modelled as unknown had a multivariate normal prior distribution derived from
eQuilibrator [@beberEQuilibratorPlatformEstimation2021].

The values for $\Delta_fG$ parameters, as well as all other model parameters,
can be found by inspecting the file `priors.toml` which is online at <https://github.com/biosustain/Methionine_model/blob/main/data/methionine/priors.toml>.

## Computation

We conducted adaptive Hamiltonian Monte Carlo sampling for the full and
missing-data datasets. For the full dataset we obtained 1000 post-warmup
samples each from 4 independent Markov chains after 1000 warm-up samples and
“hot-starting” with a mass metric output by a previous model run.

For the missing-data dataset 250 post-warmup samples were taken from 4
indpendent Markov  chains after 100 warmup samples. The sampling was initialised
using the mass matric from the complete measurement dataset and the warmup 
consisted of step size adaption for 100 samples. The resulting posterior distribution
had an $\hat{R} = 1.01$ for the log-probability and did not exhibit post-warmup 
divergences that were not a result of differential equation errors.

# Laplace approximation case study

To compare MCMC sampling with Laplace approximation we used a different model
with fewer parameters and state variables. This model was chosen because
we were not able to generate results for our methionine model using Laplace
approximation. The simpler case still serves to illustrate the general issues
with approximating the posterior distributions of Bayesian kinetic models using
the Laplace method, and that the associated numerical instability is another
reason to prefer other methods where possible.

The full Maud input folders used for our Laplace approximation case study can
be found at <https://github.com/biosustain/Methionine_model/tree/main/data/example_ode>
and <https://github.com/biosustain/Methionine_model/tree/main/data/example_ode_laplace>.

To generate Laplace samples we used Maud's Laplace mode.

# Multimodal posterior distributions

<!-- :::{.callout-important title="Changes"} -->
While we have not yet observed this in practice, we expect that Maud is capable
of accurately sampling mildly multi-modal posterior distributions, i.e. those
for which regions of parameter space with high probability density are not
very sharply separated. This is because this ability depends primarily on the
underlying inference algorithm, and adaptive Hamiltonian Monte Carlo is known
to be able to sample many such posterior distributions. See, for example,
@mangiolaSccompRobustDifferential2023, which reports the successful use of
adaptive Hamiltonian Monte Carlo to sample posterior distributions involving
mixtures of Gaussian distributions, which are typically multi-modal.

Moreover, in cases where it fails to sample a multi-modal posterior distribution
adaptive Hamiltonian Monte Carlo typically exhibits poor mixing and divergent
transitions, which Maud is set up to detect automatically, making this case easy
for users to diagnose.

Nonetheless, it is also well known that strongly multi-modal posterior
distributions pose a problem for adaptive Hamiltonian Monte Carlo, as well as
other MCMC algorithms. See [@betancourtmichaelMarkovChainMonte2020, §3.2] for
a general discussion of this issue. @mangoubiDoesHamiltonianMonte2018 explores
the related question of the relative performance of Hamiltonian Monte Carlo vs
random walk Metropolis Hastings, finding that, for certain cases where the target
distribution exhibits highly disconnected and similarly shaped modes, the two
algorithms have similar performance.

It is therefore important to pay careful attention to MCMC diagnostics and
monitor developments in computational statistics that might expand the range of
posterior distributions that can practically be sampled.

@zamora-silleroEfficientCharacterizationHighdimensional2011 introduces
the method HYPERSPACE, which is designed to identify disconnected regions
of a biochemical parameter space and characterise them by estimating
their dimensions. HYPERSPACE does not perform parameter inference and is
therefore not directly comparable with algorithms like adaptive Hamiltonian
Monte Carlo as used in Maud, but could plausibly be adapted into such an
algorithm, for example by supplementation with importance sampling (see
@vehtariParetoSmoothedImportance2022). However, any such algorithm would
be limited by HYPERSPACE's dependency on a random walk Metropolis Hastings
algorithm. This kind of algorithm's performance is known to scale poorly
compared with Hamiltonian Monte Carlo algorithms as the number of parameters
increases: see @mangoubiDoesHamiltonianMonte2018. Since Maud aims to fit models
with hundreds of parameters, this means that it is unlikely that HYPERSPACE can
directly be used to improve Maud's efficiency.
<!-- ::: -->

# References

::: {#refs}
:::